---
title: "BNST 2025 Analysis: Cell Type Assignment"
author: "Sinead Archdeacon"
date: "2026-03-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,  
                      warning = FALSE)
```

## Introduction

The next step of this snRNA-seq analysis is cell type assignment, which was done through cell type label transfer from reference data set from the Allen Brain Atlas. 
<https://alleninstitute.github.io/abc_atlas_access/descriptions/WMB-10Xv3.html>

*Note:* I kept control and stress data separate for this step.

```{r Environment setup}

# Increase capacity of R: necessary at the start of every session
options(future.globals.maxSize = 60000 * 1024^2)

# Load libraries
library(Seurat)
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(sctransform) 
library(harmony)
library(zellkonverter)
library(SingleCellExperiment)
library(AnnotationDbi)
library(org.Mm.eg.db)
library(cowplot)
```

## Reading in reference data

First, read in the raw data files.

```{r Read in reference data}

# Download reference dataset files into working directory
options(timeout = max(600, getOption("timeout")))  # 10 minute timeout

# reference expression matrix
download.file(
  url = "https://allen-brain-cell-atlas.s3.us-west-2.amazonaws.com/expression_matrices/WMB-10Xv3/20230630/WMB-10Xv3-PAL-raw.h5ad",
  destfile = "WMB-10Xv3-PAL-raw.h5ad",
  mode = "wb",
  method = "curl" # more stable for large file downloads
)

# reference metadata
download.file(
  url = "https://allen-brain-cell-atlas.s3-us-west-2.amazonaws.com/metadata/WMB-10X/20230630/cell_metadata.csv",
  destfile = "WMB-10X-cell_metadata.csv", # this contains the cluster alias
  mode = "wb",
  method = "curl"
)
download.file(
  url = "https://allen-brain-cell-atlas.s3-us-west-2.amazonaws.com/metadata/WMB-10X/20230630/region_of_interest_metadata.csv",
  destfile = "WMB-10X-roi_metadata.csv",
  mode = "wb",
  method = "curl"
)
download.file(
  url = "https://allen-brain-cell-atlas.s3-us-west-2.amazonaws.com/metadata/WMB-taxonomy/20230630/cluster_to_cluster_annotation_membership.csv",
  destfile = "WMB-10X-celltypes_metadata.csv",
  mode = "wb",
  method = "curl"
)
```

Then convert the reference data to a Seurat object with metadata.

```{r Seurat object creation}

sce <- readH5AD("WMB-10Xv3-PAL-raw.h5ad", use_hdf5 = TRUE, reader = "R")
sce
colnames(colData(sce))
head(rownames(sce)) # should be ENSMBL IDS
gene_mapping <- mapIds(
  org.Mm.eg.db,
  keys = rownames(sce),
  column = "SYMBOL",
  keytype = "ENSEMBL",
  multiVals = "first"
)
sce <- sce[!is.na(gene_mapping), ]
rownames(sce) <- gene_mapping[!is.na(gene_mapping)]
assay(sce, "X") <- as.matrix(assay(sce, "X"))
allen_ref <- as.Seurat(sce, counts = "X", data = NULL)   
head(rownames(allen_ref)) # should be gene symbols    

# Add metadata 
cell_metadata <- read.csv("WMB-10X-cell_metadata.csv", stringsAsFactors = FALSE) 
roi_metadata  <- read.csv("region_of_interest_metadata.csv", stringsAsFactors = FALSE)
merged_metadata <- merge(cell_metadata, roi_metadata,
                         by.x = "region_of_interest_acronym", by.y = "acronym", all.x = TRUE)
rownames(merged_metadata) <- merged_metadata$cell_label
merged_metadata <- merged_metadata[colnames(allen_ref), ]
allen_ref <- AddMetaData(allen_ref, metadata = merged_metadata)

# Match "cluster_alias" to "cluster annotation term"
cluster_alias_to_cluster_annotation_term_name <- read.csv("WMB-10X-celltypes_metadata.csv") 

# Reformat cluster_alias_to_cluster_annotation_term table to enable adding multiple cell type granularities (e.g. division, class, subclass)
cluster_alias_to_cluster_annotation_term_name_wide <- 
  cluster_alias_to_cluster_annotation_term_name %>%
  dplyr::select(
    cluster_alias,
    cluster_annotation_term_set_name,
    cluster_annotation_term_name) %>%
  tidyr::pivot_wider(
    names_from = cluster_annotation_term_set_name,
    values_from = cluster_annotation_term_name)

# Add cluster annotation term 
allen_ref@meta.data <- allen_ref@meta.data %>%
  rownames_to_column(var = "temp_cell_barcode") %>%
  left_join(cluster_alias_to_cluster_annotation_term_name_wide, by = "cluster_alias") %>%
  column_to_rownames(var = "temp_cell_barcode")

# Ensure all new columns are plain characters (no names)
allen_ref$division <- as.character(allen_ref$division)
allen_ref$class    <- as.character(allen_ref$class)
allen_ref$subclass <- as.character(allen_ref$subclass)
allen_ref$supertype<- as.character(allen_ref$supertype)
allen_ref$cluster  <- as.character(allen_ref$cluster)

# Inspect metadata
length(unique(allen_ref$division)) #7, most broad
length(unique(allen_ref$class)) #22
length(unique(allen_ref$subclass)) #133, this resolution was not needed for my data
length(unique(allen_ref$supertype)) #424, this resolution was not needed for my data
length(unique(allen_ref$cluster)) #1251, most granular, this resolution was not needed for my data

# Remove irrelevant metadata columns 
allen_ref@meta.data <- allen_ref@meta.data[ , !(colnames(allen_ref@meta.data) %in% 
  c("anatomical_division_label", "region_of_interest_acronym", "library_method", "dataset_label", 
  "matrix_label", "label", "name", "order", "color_hex_triplet"))]
```

## Reference data preprocessing

Follows the same steps as the BNST data preprocessing.

``` {r Preprocess the reference data}

allen_ref <- NormalizeData(allen_ref)
allen_ref <- FindVariableFeatures(allen_ref, selection.method = "vst", nfeatures = 3000)
allen_ref <- ScaleData(allen_ref)
allen_ref <- RunPCA(allen_ref)
ElbowPlot(allen_ref, ndims= 50) # pick a dim based on the elbow plot
allen_ref <- FindNeighbors(allen_ref, dims = 1:50) # clusters already provided
set.seed(123)
allen_ref <- RunUMAP(allen_ref, dims = 1:50, return.model = TRUE)

DimPlot(allen_ref, reduction = "umap", group.by = "division", label = FALSE) + 
  ggtitle("Reference Dataset UMAP by Division") + theme(legend.position = "right")
DimPlot(allen_ref, reduction = "umap", group.by = "class", label = FALSE) + 
  ggtitle("Reference Dataset UMAP by Class") + theme(legend.position = "right")

# Clean up environment 
rm(cell_metadata, cluster_alias_to_cluster_term_name, cluster_alias_to_cluster_term_name_wide, merged_metadata, roi_metadata, sce)

allen_ref <- saveRDS(allen_ref, file = "C:/shares/btrainor/Sinead/GitHub/allen_ref_PAL.rds")
``` 

## Controls label transfer

First, read in the pre-processed query data and recompute the RNA assay for comparison to the reference data (which only has an RNA assay, not SCT).

```{r CONTROLS: read in pre-processed query data}
BNST.controls <- readRDS(file = "C:/shares/btrainor/Sinead/GitHub/BNST.controls.rds")
DefaultAssay(BNST.controls) <- "RNA"
BNST.controls <- NormalizeData(BNST.controls, verbose = FALSE)
BNST.controls <- FindVariableFeatures(BNST.controls, nfeatures = 3000, verbose = FALSE)
BNST.controls <- ScaleData(BNST.controls, verbose = FALSE)
BNST.controls <- RunPCA(BNST.controls, verbose = FALSE)
```

### Data mapping
This uses a unimodal UMAP projection. **MapQuery** is recommended instead of TransferData for Seurat v4.3+

*See output in BNST.controls_celltype_mapping_plots.pdf*

```{r CONTROLS: Map query data to reference data}

anchors <- FindTransferAnchors(reference = allen_ref, query = BNST.controls, normalization.method = "LogNormalize",
                               dims = 1:50, reference.reduction = "pca")
# 8859 anchors found
BNST.controls <- MapQuery(anchorset = anchors, reference = allen_ref, query = BNST.controls,
                                  refdata = list(division = "division", class = "class"), 
                                  reference.reduction = "pca", reduction.model = "umap")

# Visualize query cells alongside reference
p1a <- DimPlot(allen_ref, reduction = "umap", group.by = "division", label = TRUE, label.size = 3,
              repel = TRUE) + NoLegend() + ggtitle("Reference Divisions")
p1b <- DimPlot(BNST.controls, reduction = "ref.umap", group.by = "predicted.division", label = TRUE,
              label.size = 3, repel = TRUE) + NoLegend() + ggtitle("BNST Controls Divisions")
print(p1a + p1b)
p2a <- DimPlot(allen_ref, reduction = "umap", group.by = "class", label = TRUE, label.size = 3,
              repel = TRUE) + ggtitle("Reference Classes") + theme(legend.position = "right")
p2b <- DimPlot(BNST.controls, reduction = "ref.umap", group.by = "predicted.class", label = TRUE,
              label.size = 3, repel = TRUE) + ggtitle("BNST Controls Classes") + theme(legend.position = "right")
print(p2a + p2b)

# More visualizing
## Create UMAP plots for each Seurat object and extract UMAP coordinates
umap_coords1 <- as.data.frame(Embeddings(allen_ref, reduction = "umap"))   
umap_coords2 <- as.data.frame(Embeddings(BNST.controls, reduction = "ref.umap"))
colnames(umap_coords2) <- c("umap_1", "umap_2")
umap_coords1$origin = "Reference Data"
umap_coords2$origin = "BNST Controls Data"

## Merge the UMAP coordinates based on common identifier
merged_umap_coords <- rbind(umap_coords1, umap_coords2)
merged_umap_coords$origin <- factor(merged_umap_coords$origin, levels = c("BNST Controls Data","Reference Data"))
merged_umap_coords <- merged_umap_coords %>% arrange(desc(origin))

## Plot the overlay of UMAPs
p3 <- ggplot(data = merged_umap_coords, aes(x = umap_1, y = umap_2, color = origin)) +
  geom_point(size = 0.5) +
  scale_color_manual(values = c("red", "black")) +
  theme_cowplot()
print(p3)
``` 

### By-cluster voting: division
Cluster identities are assigned by selecting the transferred division label that is most common in that cluster's cells.

*See output in BNST.controls_by-cluster_voting_plots.pdf*

Cluster purity refers to the number of cells in cluster that fall under most popular division, divided by the total
number of cells. The purity should be above 60% for most clusters.

```{r CONTROLS: By-cluster voting: division}

print(unique(BNST.controls$predicted.division)) 
```

There were 7 unique divisions assigned through data mapping, and I tried to find a query resolution that gave a similar number of clusters.

```{r CONTROLS: clustering for division, results = 'hide'}
# try a range of clusters
res.vec <- c(0.01, 0.03, 0.05, 0.07)
BNST.controls <- FindClusters(BNST.controls, graph.name = "SCT_snn", resolution = res.vec)
```

res.0.05 gave 10 clusters, so I picked this.

```{r CONTROLS: table, histogram}

# make table with number of cells with each predicted division for each cluster
clust_count_table_division <- BNST.controls@meta.data %>% group_by(SCT_snn_res.0.05, predicted.division) %>% dplyr::count()   
pct_df <- clust_count_table_division %>%
  group_by(SCT_snn_res.0.05) %>%
  mutate(max_n = max(n)) %>%   # find number of cells belonging to most popular division in each cluster
  ungroup() %>%
  group_by(SCT_snn_res.0.05, max_n) %>%
  summarise(percentage = max(n) / sum(n) * 100) %>%   # compute cluster purity 
  ungroup()

hist(
  pct_df$percentage,
  main = "Histogram of Cluster Purity (res.0.05)",
  xlab = "Purity (% of cells in dominant predicted division)",
  ylab = "Number of clusters"
)
```

The histogram shows high cluster purity at this resolution.

```{r CONTROLS: by-cluster voting for division continued}

# table of most popular division for each cluster
max_table <- clust_count_table_division %>% group_by(SCT_snn_res.0.05) %>% slice_max(n) %>% ungroup()  
max_map <- setNames(max_table$predicted.division, nm=max_table$SCT_snn_res.0.05)

# assign the most popular division to each cluster
BNST.controls@meta.data$pred.div.res.0.05 <- max_map[BNST.controls@meta.data$SCT_snn_res.0.05]
BNST.controls@meta.data$pred.div.res.0.05 <- gsub(" ", "_" ,BNST.controls@meta.data$pred.div.res.0.05)

# visualize
p4 <- DimPlot(BNST.controls, reduction = "umap", group.by = "pred.div.res.0.05", label = TRUE,
              pt.size=1, label.size = 3, repel = TRUE, label.box = TRUE) + NoLegend() + 
  ggtitle("BNST Controls Transferred Labels by Division")
print(p4)
```  

### By-cluster voting: class
As with divisions, cluster identities are assigned by selecting the transferred class label most common in that cluster's cells.

```{r CONTROLS By-cluster voting: class}

print(unique(BNST.controls$predicted.class))
```

There were 22 unique classes assigned through data mapping. As with divisions, I wanted to pick a query resolution that gave a similar number of clusters. 

```{r CONTROLS: clustering for division, results = 'hide'}
# try a range of clusters
res.vec <- c(0.1, 0.2, 0.3, 0.4, 0.5)
BNST.controls <- FindClusters(BNST.controls, graph.name = "SCT_snn", resolution = res.vec)
```

res.0.3 gives 24 clusters, so I picked this.

```{r CONTROLS: table, histogram}

# make table with number of cells with each predicted class for each cluster
clust_count_table_class <- BNST.controls@meta.data %>% group_by(SCT_snn_res.0.3, predicted.class) %>%
dplyr::count()  
pct_df <- clust_count_table_class %>%
  group_by(SCT_snn_res.0.3) %>%
  mutate(max_n = max(n)) %>%   # find number of cells belonging to most popular class in each cluster
  ungroup() %>%
  group_by(SCT_snn_res.0.3, max_n) %>%
  summarise(percentage = max(n) / sum(n) * 100) %>%   # compute cluster purity
  ungroup()

hist(
  pct_df$percentage,
  main = "Histogram of Cluster Purity (res.0.3)",
  xlab = "Purity (% of cells in dominant predicted class)",
  ylab = "Number of clusters"
)
```

The histogram shows that most clusters have high purity at this resolution. 

```{r CONTROLS: by-cluster voting for class continued}

# table of most popular class for each cluster
max_table <- clust_count_table_class %>% group_by(SCT_snn_res.0.3) %>% slice_max(n) %>% ungroup()  
max_map <- setNames(max_table$predicted.class, nm=max_table$SCT_snn_res.0.3)

# assign most popular class to each cluster
BNST.controls@meta.data$pred.class.res.0.3 <- max_map[BNST.controls@meta.data$SCT_snn_res.0.3]
BNST.controls@meta.data$pred.class.res.0.3 <- gsub(" ", "_" ,BNST.controls@meta.data$pred.class.res.0.3)

# visualize 
p5 <- DimPlot(BNST.controls, reduction = "umap", group.by = "pred.class.res.0.3", label = TRUE,
              pt.size=1, label.size = 3, repel = TRUE, label.box = TRUE) + NoLegend() + 
  ggtitle("BNST Controls Transferred Labels by Class")
print(p5)
``` 

### Save progress

```{r CONTROLS: Save progress}

# Remove irrelevant metadata columns
BNST.controls@meta.data <- BNST.controls@meta.data[ , !(colnames(BNST.controls@meta.data) %in% 
            c("condition", "DoubletStatus", "SCT_snn_res.0.01", "SCT_snn_res.0.02", "SCT_snn_res.0.04",
            "SCT_snn_res.0.05", "SCT_snn_res.0.06", "SCT_snn_res.0.07", "SCT_snn_res.0.1", "SCT_snn_res.0.2",  
            "SCT_snn_res.0.3", "SCT_snn_res.0.4", "SCT_snn_res.0.5", "predicted.division.score", 
            "predicted.class.score"))]

# Clean up recursions before saving  #NOT WORKING
BNST.controls@tools <- list()
BNST.controls@anchors <- NULL
if ("ref.umap" %in% names(BNST.controls@reductions)) {
  BNST.controls@reductions$ref.umap@misc$model <- NULL
}
if ("umap" %in% names(BNST.controls@reductions)) {
  BNST.controls@reductions$umap@misc$model <- NULL
}
BNST.controls@reductions$ref.umap@misc <- list()
BNST.controls@reductions$umap@misc <- list()
BNST.controls@misc <- list()

BNST.controls.clean <- DietSeurat(
  BNST.controls,
  counts = TRUE,
  data = TRUE,
  scale.data = FALSE,
  dimreducs = c("pca", "umap", "ref.umap"),
  assays = c("RNA", "SCT"))

BNST.controls[["ref.umap"]] <- NULL
BNST.controls[["umap"]] <- NULL
BNST.controls@tools <- NULL

saveRDS(BNST.controls, file = "C:/shares/btrainor/Sinead/GitHub/BNST.controls_mapped.rds")

``` 

## Stress label transfer
This uses the same workflow as for the control data.

```{r STRESS: Read in pre-processed query data}
BNST.stress <- readRDS(file = "C:/shares/btrainor/Sinead/GitHub/BNST.stress.rds")

# Recompute RNA assay
DefaultAssay(BNST.stress) <- "RNA"
BNST.stress <- NormalizeData(BNST.stress, verbose = FALSE)
BNST.stress <- FindVariableFeatures(BNST.stress, nfeatures = 3000, verbose = FALSE)
BNST.stress <- ScaleData(BNST.stress, verbose = FALSE)
BNST.stress <- RunPCA(BNST.stress, verbose = FALSE)
```

*See output in BNST.stress_celltype_mapping_plots.pdf*

```{r STRESS: Map query data to reference data}

anchors <- FindTransferAnchors(reference = allen_ref, query = BNST.stress, normalization.method = "LogNormalize", 
                               dims = 1:50, reference.reduction = "pca") 
# 8012 anchors found
BNST.stress <- MapQuery(anchorset = anchors, reference = allen_ref, query = BNST.stress, 
                        refdata = list(division = "division", class = "class"),
                        reference.reduction = "pca", reduction.model = "umap")

# Visualize query cells alongside reference 
p6a <- DimPlot(allen_ref, reduction = "umap", group.by = "division", label = TRUE, label.size = 3,
               repel = TRUE) + NoLegend() + ggtitle("Reference Divisions")
p6b <- DimPlot(BNST.stress, reduction = "ref.umap", group.by = "predicted.division", label = TRUE,
               label.size = 3, repel = TRUE) + NoLegend() + ggtitle("BNST Stress Divisions")
print(p6a + p6b)
p7a <- DimPlot(allen_ref, reduction = "umap", group.by = "class", label = TRUE, label.size = 3,
               repel = TRUE) + NoLegend() + ggtitle("Reference Classes")
p7b <- DimPlot(BNST.stress, reduction = "ref.umap", group.by = "predicted.class", label = TRUE,
               label.size = 3, repel = TRUE) + NoLegend() + ggtitle("BNST Stress Classes")
print(p7a + p7b)

# More visualizing
## Create UMAP plots for each Seurat object and extract UMAP coordinates
umap_coords1 <- as.data.frame(Embeddings(allen_ref, reduction = "umap"))  
umap_coords2 <- as.data.frame(Embeddings(BNST.stress, reduction = "ref.umap"))
colnames(umap_coords2) <- c("umap_1", "umap_2")
umap_coords1$origin = "Reference Data"
umap_coords2$origin = "BNST Stress Data"

## Merge the UMAP coordinates based on common identifier
merged_umap_coords <- rbind(umap_coords1, umap_coords2)
merged_umap_coords$origin <- factor(merged_umap_coords$origin, levels = c("BNST Stress Data","Reference Data"))
merged_umap_coords <- merged_umap_coords %>% arrange(desc(origin))

## Plot the overlay of UMAPs
p8 <- ggplot(data = merged_umap_coords, aes(x = umap_1, y = umap_2, color = origin)) +
  geom_point(size = 0.5) +
  scale_color_manual(values = c("red", "black")) +
  theme_cowplot()
print(p8)

```

### By-cluster voting: division
This uses the same approach as for the control data.

*See output in BNST.stress_by-cluster_voting_plots.pdf*

```{r STRESS: By-cluster voting: division}

print(unique(BNST.stress$predicted.division)) 
```

There were 7 unique divisions assigned through data mapping.

```{r STRESS: clustering for division, results = 'hide'}

# try a range of clustering resolutions
res.vec <- c(0.01, 0.02, 0.03, 0.04, 0.05)
BNST.stress <- FindClusters(BNST.stress, graph.name = "SCT_snn", resolution = res.vec)
```

res0.02 gave 7 clusters, so I used this. 

```{r STRESS: table, histogram}

# make table with number of cells with each predicted divisions for each cluster
clust_count_table_division <- BNST.stress@meta.data %>% group_by(SCT_snn_res.0.02, predicted.division) %>% dplyr::count()
pct_df <- clust_count_table_division %>%
  group_by(SCT_snn_res.0.02) %>%
  mutate(max_n = max(n)) %>%   # find number of cells belonging to most popular division in each cluster
  ungroup() %>%
  group_by(SCT_snn_res.0.02, max_n) %>%
  summarise(percentage = max(n) / sum(n) * 100) %>%   # compute cluster purity 
  ungroup()

hist(
  pct_df$percentage,
  main = "Histogram of Cluster Purity (res.0.02)",
  xlab = "Purity (% of cells in dominant predicted division)",
  ylab = "Number of clusters"
)
```

The histogram shows high cluster purity at this resolution.

```{r STRESS: by-cluster voting for division continued}

# table of most popular division for each cluster
max_table <- clust_count_table_division %>% group_by(SCT_snn_res.0.02) %>% slice_max(n) %>% ungroup()  
max_map <- setNames(max_table$predicted.division, nm=max_table$SCT_snn_res.0.02)

# assign most popular division to each cluster
BNST.stress@meta.data$pred.div.res.0.02 <- max_map[BNST.stress@meta.data$SCT_snn_res.0.02]
BNST.stress@meta.data$pred.div.res.0.02 <- gsub(" ", "_" ,BNST.stress@meta.data$pred.div.res.0.02)

# visualize 
p9 <- DimPlot(BNST.stress, reduction = "umap", group.by = "pred.div.res.0.02", label = TRUE,
             pt.size=1, label.size = 3, repel = TRUE, label.box = TRUE) + NoLegend() + 
  ggtitle("BNST stress Transferred Labels by Division, res.0.02")
print(p9)
# only 5 of the 7 divisions apply
```

### By-cluster voting: class

```{r STRESS: By-cluster voting: class}

print(unique(BNST.stress$predicted.class)) 
```

There were 17 unique classes assigned through data mapping.

```{r STRESS: clustering for class, results = 'hide'}

# try a range of clustering resolutions
res.vec <- c(0.1, 0.2, 0.3, 0.4, 0.5)
BNST.stress <- FindClusters(BNST.stress, graph.name = "SCT_snn", resolution = res.vec)
```

res.0.1 gives 17 clusters, so I picked this.

```{r STRESS: table, histogram}

# make table with number of cells with each predicted class for each cluster
clust_count_table_class <- BNST.stress@meta.data %>% group_by(SCT_snn_res.0.1, predicted.class) %>% dplyr::count()
pct_df <- clust_count_table_class %>%
  group_by(SCT_snn_res.0.1) %>%
  mutate(max_n = max(n)) %>%   # find number of cells belonging to most popular class in each cluster
  ungroup() %>%
  group_by(SCT_snn_res.0.1, max_n) %>%
  summarise(percentage = max(n) / sum(n) * 100) %>%   # compute cluster purity 
  ungroup()

hist(
  pct_df$percentage,
  main = "Histogram of Cluster Purity (res0.1)",
  xlab = "Purity (% of cells in dominant predicted class)",
  ylab = "Number of clusters"
)
```

Clusters 2, 6, 9, 10, 11, 12, 14, and 15 have <60% purity, so I inspected these clusters to confirm that they didn't need to be re-clustered to add more information. If a cluster was somewhat evenly split between neuronal and non-neuronal cells, I reclustered it to help me identify neurons more accurately.

```{r STRESS: inspect low-purity clusters}

cluster2 <- subset(BNST.stress, SCT_snn_res.0.1 %in% "2") 
table(cluster2$predicted.class) # almost all are either CNU-HYa GABA or just CNU GABA --> don't need to recluster

cluster6 <- subset(BNST.stress, SCT_snn_res.0.1 %in% "6") 
table(cluster6$predicted.class) 
# out of 4000 cells 2300 are oligo and 850 are astro, about 670 are GABA neurons. So 80% are glial --> don't recluster

cluster9 <- subset(BNST.stress, SCT_snn_res.0.1 %in% "9") 
table(cluster9$predicted.class) # 50% CNU-HYa GABA, 27% immune --> recluster

cluster10 <- subset(BNST.stress, SCT_snn_res.0.1 %in% "10") 
table(cluster10$predicted.class) # almost all are either CNU-HYa GABA or just CNU GABA --> don't recluster

cluster11 <- subset(BNST.stress, SCT_snn_res.0.1 %in% "11") 
table(cluster11$predicted.class) # 50% CNU-HYa GABA, 40% astro --> recluster

cluster14 <- subset(BNST.stress, SCT_snn_res.0.1 %in% "14") 
table(cluster14$predicted.class) # 50% CNU-HYa GABA, 35% vascular --> recluster

cluster15 <- subset(BNST.stress, SCT_snn_res.0.1 %in% "15") 
table(cluster15$predicted.class) # almost all are either CNU-HYa GABA (252) or just CNU GABA (159) --> don't recluster
```

```{r STRESS: recluster clusters 9, 11, 14}

cells_to_refine <- WhichCells(BNST.stress, expression = SCT_snn_res.0.1 %in% c("9", "11", "14"))
clusters_to_refine <- subset(BNST.stress, cells = cells_to_refine)
colnames(clusters_to_refine@meta.data)[colnames(clusters_to_refine@meta.data) == "SCT_snn_res.0.1"] <- "parent_clustering"

DefaultAssay(clusters_to_refine) <- "SCT"
clusters_to_refine <- RunPCA(clusters_to_refine, features = VariableFeatures(BNST.stress_nonfilt))
clusters_to_refine <- RunHarmony(clusters_to_refine, assay.use = "SCT", group.by.vars = "orig.ident")
ElbowPlot(clusters_to_refine, ndims = 50)

refined_list <- lapply(unique(clusters_to_refine$parent_clustering), function(parent_cluster){
  sub_cells <- WhichCells(clusters_to_refine, expression = parent_clustering == parent_cluster)
  sub_obj <- subset(clusters_to_refine, cells = sub_cells)
  sub_obj <- RunPCA(sub_obj, features = VariableFeatures(sub_obj))
  sub_obj <- FindNeighbors(sub_obj, dims = 1:10)   # use top 10 PCs for local structure
  sub_obj <- FindClusters(sub_obj, resolution = 0.1)
  sub_obj$refined_clustering <- paste0(parent_cluster, "_sub", Idents(sub_obj))
  refined_vec <- sub_obj$refined_clustering  # return as named vector (cell names preserved)
  names(refined_vec) <- colnames(sub_obj)
  return(refined_vec)
})
refined_all <- unlist(refined_list)
clusters_to_refine$refined_clustering <- refined_all[colnames(clusters_to_refine)]

# plot UMAPs to check
cluster11 <- subset(clusters_to_refine, parent_clustering == "11")
cluster11$refined_clustering <- factor(cluster11$refined_clustering)
DimPlot(cluster11, reduction = "umap", group.by = "refined_clustering", label = FALSE) # 4 subclusters, look decent

cluster9 <- subset(clusters_to_refine, parent_clustering == "9")
cluster9$refined_clustering <- factor(cluster9$refined_clustering)
DimPlot(cluster9, reduction = "umap", group.by = "refined_clustering", label = FALSE) # 5 subclusters, look ok

cluster14 <- subset(clusters_to_refine, parent_clustering == "14")
cluster14$refined_clustering <- factor(cluster14$refined_clustering)
DimPlot(cluster14, reduction = "umap", group.by = "refined_clustering", label = FALSE) # 2 subclusters, look very mixed

# create metadata column in parent object for subclusters
BNST.stress$refined_subcluster <- NA
BNST.stress$refined_subcluster[cells_to_refine] <- clusters_to_refine$refined_clustering
BNST.stress$final_cluster_label <- ifelse(
  is.na(BNST.stress$refined_subcluster),
  as.character(BNST.stress$SCT_snn_res.0.1),
  as.character(BNST.stress$refined_subcluster)
)
```

```{r STRESS: table, histogram}

clust_count_table_class2 <- BNST.stress@meta.data %>% group_by(final_cluster_label, predicted.class) %>% dplyr::count() 
pct_df2 <- clust_count_table_class3 %>%
  group_by(final_cluster_label) %>%
  mutate(max_n = max(n)) %>%   # find number of cells belonging to most popular class in each cluster
  ungroup() %>%
  group_by(final_cluster_label, max_n) %>%
  summarise(percentage = max(n) / sum(n) * 100) %>%   # compute cluster purity 
  ungroup()

hist(
  pct_df2$percentage,
  main = "Histogram of Cluster Purity (refined)",
  xlab = "Purity (% of cells in dominant predicted class)",
  ylab = "Number of clusters"
)
```

Clusters with <60% purity: 2 (ok), 6 (ok), 9sub0, 9sub1, 10 (ok), 12 (ok), 15 (ok)

9sub0 has 1212 cells, 9sub1 has 443 --> less than 2% of entire data, so ok to throw out. Also, the most common label is Oligo/Immune, which I'm not analyzing here anyways.

```{r throw out cluster 9}

BNST.stress_nonfilt <- subset(BNST.stress_nonfilt, final_cluster_label %in% c("9_sub0", "9_sub1"), invert = TRUE) 
# don't need to renormalize since removing so few cells
```

```{r STRESS: by-cluster voting for class continued}

max_map <- setNames(max_table$predicted.class, nm=max_table$final_cluster_label)

# assign most popular class to each cluster
BNST.stress@meta.data$pred.class.finalized <- max_map[BNST.stress@meta.data$final_cluster_label]
BNST.stress@meta.data$pred.class.finalized <- gsub(" ", "_" ,BNST.stress@meta.data$pred.class.finalized)

BNST.stress$pred.class.finalized[BNST.stress$SCT_snn_res.0.1 %in% "6"] <- "glial"
BNST.stress$pred.class.finalized[BNST.stress$SCT_snn_res.0.1 %in% "12"] <- "ambiguous GABA/glut"

# visualize 
p10 <- DimPlot(BNST.stress, reduction = "umap", group.by = "pred.class.finalized", label = TRUE,
              pt.size=1, label.size = 3, repel = TRUE, label.box = TRUE) + NoLegend() + 
  ggtitle("BNST Stress Transferred Labels by Class")
print(p10)
```

### Save progress

``` {r STRESS: Save progress}

# remove irrelevant metadata columns
BNST.stress@meta.data <- BNST.stress@meta.data[ , !(colnames(BNST.stress@meta.data) %in% 
      c("condition", "DoubletStatus", "SCT_snn_res.0.01", "SCT_snn_res.0.02", 
      "SCT_snn_res.0.04", "SCT_snn_res.0.05", "SCT_snn_res.0.06", "SCT_snn_res.0.08", "SCT_snn_res.0.1", 
      "predicted.division.score", "predicted.class.score"))]

# clean up recursions before saving
BNST.stress@tools <- list()
BNST.stress@anchors <- NULL
if ("ref.umap" %in% names(BNST.stress@reductions)) {
  BNST.stress=@reductions$ref.umap@misc$model <- NULL
}
if ("umap" %in% names(BNST.stress@reductions)) {
  BNST.stress@reductions$umap@misc$model <- NULL
}
BNST.stress@reductions$ref.umap@misc <- list()
BNST.stress@reductions$umap@misc <- list()
BNST.stress@misc <- list()

saveRDS(BNST.stress, file = "C:/shares/btrainor/Sinead/GitHub/BNST.stress_mapped.rds")
```
