---
title: "BNST 2025 Analysis: BNST-PVN Integration"
author: "Sinead Archdeacon"
date: "2026-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This final section compares subclusters and magnocellular/parvocellular marker expression for BNST Oxt neurons and PVN Oxt neurons (from a separate data collection). This forms the bulk of the work I presented at SfN in 2025. 

```{r Environment setup}

# Increase capacity of R: necessary at the start of every session
options(future.globals.maxSize = 60000 * 1024^2)

# Load libraries
library(Seurat)
library(dplyr)
library(ggplot2)
library(sctransform) 
library(harmony)
```

## Read in the data

```{r Read in data}

BNST.oxt_neurons <- readRDS(file = "C:/shares/btrainor/Sinead/GitHub/BNST.oxt_neurons.rds")
PVN.oxt_neurons <- readRDS(file = "C:/shares/btrainor/Sequencingdata/Audrey's code 2025/PVN_OxtNeuronSubset.rds")

BNST.oxt_neurons$region <- "BNST"
PVN.oxt_neurons$region <- "PVN"

BNST.oxt_neurons$region <- "BNST"
PVN.oxt_neurons$region <- "PVN"
table(BNST.oxt_neurons$orig.ident) # all samples are represented

PVN.oxt_neurons$orig.ident <- paste0("PVN_", PVN.oxt_neurons$orig.ident)
BNST.oxt_neurons$orig.ident <- paste0("BNST_", BNST.oxt_neurons$orig.ident)
```

## Reformat assays 

Reformat to make the two objects compatible for merging.

```{r Reformat}
BNST.oxt_neurons[["RNA"]]@layers$data.1 <- NULL # strip old data layers 
BNST.oxt_neurons[["RNA"]]@layers$data.2 <- NULL
BNST.oxt_neurons <- NormalizeData(BNST.oxt_neurons, assay = "RNA") # re-create data layer from all neurons

# check that BNST layer dims match
Layers(BNST.all_oxt_neurons[["RNA"]]) #counts,  data
counts_mat <- GetAssayData(BNST.all_oxt_neurons, assay = "RNA", layer = "counts") 
data_mat   <- GetAssayData(BNST.all_oxt_neurons, assay = "RNA", layer = "data")
dim(counts_mat) # 31946 features, 463 cells
dim(data_mat) # 31946 features, 463 cells

# join RNA counts and data layers for each of the 6 PVN samples
PVN.oxt_neurons <- JoinLayers(PVN.oxt_neurons, assay = "RNA", layers = "counts")
PVN.oxt_neurons <- JoinLayers(PVN.oxt_neurons, assay = "RNA", layers = "data")
rna_assay <- PVN.oxt_neurons[["RNA"]]
rna_assay@layers <- rna_assay@layers[c("counts", "data")] #remove scale.data
PVN.oxt_neurons[["RNA"]] <- rna_assay

# check that PVN layer dims match
Layers(PVN.oxt_neurons[["RNA"]]) #counts, data
counts_mat <- GetAssayData(PVN.oxt_neurons, assay = "RNA", layer = "counts")
data_mat   <- GetAssayData(PVN.oxt_neurons, assay = "RNA", layer = "data")
dim(counts_mat) # 31946 features, 1164 cells
dim(data_mat) # 31946 features, 1164 cells
all(rownames(counts_mat) == rownames(data_mat)) # TRUE

# strip original SCT assays so I can remake them using reformatted RNA assays
DefaultAssay(BNST.oxt_neurons) <- "RNA"
DefaultAssay(PVN.oxt_neurons) <- "RNA"
BNST.oxt_neurons[["SCT"]] <- NULL
PVN.oxt_neurons[["SCT"]] <- NULL
PVN.oxt_neurons[["integrated"]] <- NULL # also remove this, will run integration after merging
BNST.oxt_neurons[["prediction.score.division"]] <- NULL # remove these, irrelevant
BNST.oxt_neurons[["prediction.score.class"]] <- NULL

common_features <- intersect(rownames(BNST.oxt_neurons),
                             rownames(PVN.oxt_neurons))
BNST.oxt_neurons <- subset(BNST.oxt_neurons, features = common_features) 
PVN.oxt_neurons <- subset(PVN.oxt_neurons, features = common_features) 

# check
nrow(GetAssayData(BNST.oxt_neurons, layer = "counts")) 
nrow(GetAssayData(BNST.oxt_neurons, layer = "data")) 
nrow(GetAssayData(PVN.oxt_neurons, layer = "counts")) 
nrow(GetAssayData(PVN.oxt_neurons, layer = "data"))  

# re-order rows to be extra safe
BNST_counts <- GetAssayData(BNST.oxt_neurons, layer = "counts")
BNST_data   <- GetAssayData(BNST.oxt_neurons, layer = "data")
PVN_counts  <- GetAssayData(PVN.oxt_neurons, layer = "counts")
PVN_data    <- GetAssayData(PVN.oxt_neurons, layer = "data")
BNST.oxt_neurons[["RNA"]]@layers$counts <- BNST_counts[common_features, , drop=FALSE]
BNST.oxt_neurons[["RNA"]]@layers$data   <- BNST_data[common_features, , drop=FALSE]
PVN.oxt_neurons[["RNA"]]@layers$counts <- PVN_counts[common_features, , drop=FALSE]
PVN.oxt_neurons[["RNA"]]@layers$data   <- PVN_data[common_features, , drop=FALSE]
```

## Run SCT and clustering on PVN

```{r SCT and clustering of PVN neurons}
PVN.oxt_neurons <- SCTransform(PVN.oxt_neurons, method = "glmGamPoi", 
                                    return.only.var.genes = FALSE, 
                                    variable.features.n = 3000, 
                                    verbose = FALSE)
PVN.oxt_neurons <- RunPCA(PVN.all_oxt_neurons, verbose = FALSE)
ElbowPlot(PVN.oxt_neurons, ndims = 50)
PVN.oxt_neurons <- FindNeighbors(PVN.oxt_neurons, reduction = "pca", dims = 1:30) 
PVN.oxt_neurons <- FindClusters(PVN.oxt_neurons, graph.name = "SCT_snn", resolution = 0.05) 
PVN.oxt_neurons <- RunUMAP(PVN.oxt_neurons, reduction = "pca", dims = 1:30)
DimPlot(PVN.oxt_neurons, reduction = "umap", group.by = "SCT_snn_res.0.05") 
DimPlot(PVN.oxt_neurons, reduction = "umap", group.by = c("sex", "orig.ident", "condition"), ncol = 3) 
```

## By-cluster DE analysis of PVN Oxt neurons
*See output in just_PVN_umap.pdf,*
*PVN.oxt_neurons_top20_markers_by_cluster.csv*

```{r By-cluster DE analysis of just PVN Oxt neurons}

DefaultAssay(PVN.oxt_neurons) = "SCT"
PVN.oxt_neurons <- PrepSCTFindMarkers(PVN.oxt_neurons) 
markers_by_cluster <- FindAllMarkers(
  PVN.oxt_neurons,
  assay = "SCT",               
  layer = "data",              # use normalized counts in SCT assay
  only.pos = TRUE,             # return only upregulated genes 
  min.pct = 0.1,               # gene must be expressed in at least 10% of cells in either group
  logfc.threshold = 0.25       # minimum log2 fold-change
)

top20_markers <- markers_by_cluster %>%
  filter(p_val_adj < 0.05, !grepl("LOC", gene)) %>%
  group_by(cluster) %>%
  slice_max(order_by = avg_log2FC, n = 20) 
top20_markers <- top20_markers %>% arrange(as.numeric(cluster))

# reorder columns for clarity
top20_markers <- top20_markers[, c("cluster", "gene", setdiff(names(top20_markers), c("cluster", "gene")))] 

write.csv(top20_markers, file = "C:/shares/btrainor/Sinead/GitHub/PVN.oxt_neurons_top20_markers_by_cluster.csv", row.names = FALSE) 

PVN.oxt_neurons@meta.data$region_clusters <- NA
PVN.oxt_neurons@meta.data$region_clusters[PVN.oxt_neurons$SCT_snn_res.0.05 == "0"] <- "Oxt-Crim1" #792 cells
PVN.oxt_neurons@meta.data$region_clusters[PVN.oxt_neurons$SCT_snn_res.0.05 == "1"] <- "Oxt-Slit3" #264 cells
PVN.oxt_neurons@meta.data$region_clusters[PVN.oxt_neurons$SCT_snn_res.0.05 == "2"] <- "Oxt-Esr2" #108 cells

PVN.oxt_neurons@meta.data$region_clusters2 <- NA
PVN.oxt_neurons@meta.data$region_clusters2[PVN.oxt_neurons$SCT_snn_res.0.05 == "0"] <- "PVN_0" #792 cells
PVN.oxt_neurons@meta.data$region_clusters2[PVN.oxt_neurons$SCT_snn_res.0.05 == "1"] <- "PVN_1" #264 cells
PVN.oxt_neurons@meta.data$region_clusters2[PVN.oxt_neurons$SCT_snn_res.0.05 == "2"] <- "PVN_2" #108 cells

DimPlot(PVN.oxt_neurons, reduction = "umap", group.by = "region_clusters", label = TRUE, label.size = 5) 

p <- DimPlot(PVN.oxt_neurons, reduction = "umap",
  group.by = "region_clusters")
LabelClusters(p, id = "region_clusters", size = 5, fontface = "bold")
print(p)

# save updated object
saveRDS(PVN.oxt_neurons, file = "C:/shares/btrainor/Sinead/GitHub/PVN.oxt_neurons.rds")
```

## Run SCT and clustering on BNST

```{r SCT and clustering of BNST neurons}

BNST.oxt_neurons <- SCTransform(BNST.oxt_neurons, method = "glmGamPoi", 
                                    return.only.var.genes = FALSE, 
                                    variable.features.n = 3000, 
                                    verbose = FALSE)
BNST.oxt_neurons <- RunPCA(BNST.oxt_neurons, verbose = FALSE)
ElbowPlot(BNST.oxt_neurons, ndims = 50)
BNST.oxt_neurons <- FindNeighbors(BNST.oxt_neurons, reduction = "pca", dims = 1:5) 
BNST.oxt_neurons <- FindClusters(BNST.oxt_neurons, graph.name = "SCT_snn", resolution = 0.05) 
BNST.oxt_neurons <- RunUMAP(BNST.oxt_neurons, dims = 1:5)  
DimPlot(BNST.oxt_neurons, reduction = "umap", group.by = "SCT_snn_res.0.05") #2 clusters
DimPlot(BNST.oxt_neurons, reduction = "umap", group.by = c("sex", "orig.ident", "condition"), ncol = 3) 
```

## By-cluster DE analysis of BNST Oxt neurons
*See output in BNST.all_oxt_neurons_plots.pdf.*
*BNST.all_oxt_neurons_top20_markers_by_cluster.csv*

```{r By-cluster DE analysis of just BNST Oxt neurons}

DefaultAssay(BNST.oxt_neurons) = "SCT"
BNST.oxt_neurons <- PrepSCTFindMarkers(BNST.oxt_neurons) 
markers_by_cluster <- FindAllMarkers(
  BNST.oxt_neurons,
  assay = "SCT",               
  layer = "data",              # use normalized counts in SCT assay
  only.pos = TRUE,             # return only upregulated genes 
  min.pct = 0.1,               # gene must be expressed in at least 10% of cells in either group
  logfc.threshold = 0.25       # minimum log2 fold-change
)

top20_markers <- markers_by_cluster %>% 
  filter(p_val_adj < 0.05, !grepl("LOC", gene)) %>%
  group_by(cluster) %>%
  slice_max(order_by = avg_log2FC, n = 20) 
top20_markers <- top20_markers %>% arrange(as.numeric(cluster))

# reorder columns for clarity
top20_markers <- top20_markers[, c("cluster", "gene", setdiff(names(top20_markers), c("cluster", "gene")))] 

write.csv(top20_markers, file = "C:/shares/btrainor/Sinead/Fall2025/BNST.all_oxt_neurons_top20_markers_by_cluster.csv", row.names = FALSE) 
```

```{r continue}
BNST.oxt_neurons@meta.data$region_clusters <- NA
BNST.oxt_neurons@meta.data$region_clusters[BNST.oxt_neurons$SCT_snn_res.0.05 == "0"] <- "Oxt-Gad1"  #420 cells
BNST.oxt_neurons@meta.data$region_clusters[BNST.oxt_neurons$SCT_snn_res.0.05 == "1"] <- "Oxt-Sim1" #43 cells
BNST.oxt_neurons@meta.data$region_clusters2 <- NA
BNST.oxt_neurons@meta.data$region_clusters2[BNST.oxt_neurons$SCT_snn_res.0.05 == "0"] <- "BNST_0" #420 cells
BNST.oxt_neurons@meta.data$region_clusters2[BNST.oxt_neurons$SCT_snn_res.0.05 == "1"] <- "BNST_1" #43 cells
table(BNST.oxt_neurons$region_clusters2)
```

```{r continue}
cluster_levels <- unique(BNST.oxt_neurons$region_clusters)
BNST.oxt_neurons$region_clusters <- factor(BNST.oxt_neurons$region_clusters, levels = cluster_levels)
my_colors <- setNames(c("#9C97CC", "#66a61e"), cluster_levels) 
p <- DimPlot(BNST.oxt_neurons, reduction = "umap", group.by = "region_clusters", label = FALSE) + 
     scale_color_manual(values = my_colors) + ggtitle ("BNST OXT Neurons") + theme(legend.position = "none")
LabelClusters(p, id = "region_clusters", size = 4, fontface = "bold")
DefaultAssay(BNST.oxt_neurons) <- "RNA"
VlnPlot(BNST.oxt_neurons,
        features = c("Oxt", "Gad1", "Robo1", "Slit3", "Slc17a6", "Sim1", "Otp", "Uncx", "Pcsk1"),
        stack = TRUE, flip = TRUE) + ggtitle("BNST Oxt Neuron Markers") +
  theme(axis.text.x = element_text(angle = 0, hjust = 1), legend.position = "none") +
  xlab("Cluster") + ylab("Expression")
```

## Magno-parvo stats for BNST Oxt neuron clusters
The Kruskal-Wallis test (nonparametric ANOVA alternative) was used to check for significant differences in expression levels of magnocellular Oxt neuron markers (Foxp1, Foxp2) and parvocellular Oxt neuron markers (Adarb2, Reln, Sox5, Esr1) across BNST clusters.

```{r magno-parvo stats}

BNST.oxt_neurons <- FindNeighbors(BNST.oxt_neurons, reduction = "pca", dims = 1:5) 
BNST.oxt_neurons <- FindClusters(BNST.oxt_neurons, res = 0.05)
BNST.oxt_neurons <- RunUMAP(BNST.oxt_neurons, dims = 1:5) 

magno_genes <- c("Foxp1", "Foxp2") 
parvo_genes <- c("Adarb2", "Reln", "Sox5", "Esr1")

# use Kruskal-Wallis for counts layer
counts_df <- FetchData(BNST.oxt_neurons, vars = c("Foxp1", "SCT_snn_res.0.05"), layer = "counts")
counts_df$SCT_snn_res.0.05 <- as.factor(counts_df$SCT_snn_res.0.05)
fmla <- as.formula(paste0("Foxp1", " ~ SCT_snn_res.0.05"))
kw <- kruskal.test(fmla, data = counts_df) # statistically significant (e-22)
aggregate(Foxp1 ~ SCT_snn_res.0.05, data = counts_df, median) # higher in cluster 1

counts_df <- FetchData(BNST.oxt_neurons, vars = c("Foxp2", "SCT_snn_res.0.05"), layer = "counts")
counts_df$SCT_snn_res.0.05 <- as.factor(counts_df$SCT_snn_res.0.05)
fmla <- as.formula(paste0("Foxp2", " ~ SCT_snn_res.0.05"))
kw <- kruskal.test(fmla, data = counts_df) # statistically significant (e-26)
aggregate(Foxp2 ~ SCT_snn_res.0.05, data = counts_df, median) # higher in cluster 1

counts_df <- FetchData(BNST.oxt_neurons, vars = c("Adarb2", "SCT_snn_res.0.05"), layer = "counts")
counts_df$SCT_snn_res.0.05 <- as.factor(counts_df$SCT_snn_res.0.05)
fmla <- as.formula(paste0("Adarb2", " ~ SCT_snn_res.0.05"))
kw <- kruskal.test(fmla, data = counts_df) # statistically significant (e-15)
aggregate(Adarb2 ~ SCT_snn_res.0.05, data = counts_df, median) # higher in cluster 0

counts_df <- FetchData(BNST.oxt_neurons, vars = c("Reln", "SCT_snn_res.0.05"), layer = "counts")
counts_df$SCT_snn_res.0.05 <- as.factor(counts_df$SCT_snn_res.0.05)
fmla <- as.formula(paste0("Reln", " ~ SCT_snn_res.0.05"))
kw <- kruskal.test(fmla, data = counts_df) # statistically significant, but much less so than others (p = 0.03)
aggregate(Reln ~ SCT_snn_res.0.05, data = counts_df, median) # 0 in both clusters
aggregate(Reln ~ SCT_snn_res.0.05, data = counts_df, mean) # 1.96 in cluster 0, 1.88 in cluster 1
aggregate(Reln ~ SCT_snn_res.0.05, data = counts_df, max) # 50 in cluster 0, 32 in cluster 1

counts_df <- FetchData(BNST.oxt_neurons, vars = c("Sox5", "SCT_snn_res.0.05"), layer = "counts")
counts_df$SCT_snn_res.0.05 <- as.factor(counts_df$SCT_snn_res.0.05)
fmla <- as.formula(paste0("Sox5", " ~ SCT_snn_res.0.05"))
kw <- kruskal.test(fmla, data = counts_df) # statistically significant, but less so than others (p = 0.003)
aggregate(Sox5 ~ SCT_snn_res.0.05, data = counts_df, median) # higher in cluster 0

counts_df <- FetchData(BNST.oxt_neurons, vars = c("Esr1", "SCT_snn_res.0.05"), layer = "counts")
counts_df$SCT_snn_res.0.05 <- as.factor(counts_df$SCT_snn_res.0.05)
fmla <- as.formula(paste0("Esr1", " ~ SCT_snn_res.0.05")) 
kw <- kruskal.test(fmla, data = counts_df) # statistically significant, but much less so than others (p = 0.01)
aggregate(Esr1 ~ SCT_snn_res.0.05, data = counts_df, median) # 0 in both clusters
aggregate(Esr1 ~ SCT_snn_res.0.05, data = counts_df, mean) # 0.5 in cluster 0, 0.07 in cluster 1
aggregate(Esr1 ~ SCT_snn_res.0.05, data = counts_df, max) # 17 in cluster 0, 2 in cluster 1
```

The same approach was used to check expression in BNST Oxt neurons of the cluster markers we used for our PVN Oxt neuron data.

```{r test PVN markers}

counts_df <- FetchData(BNST.oxt_neurons, vars = c("Crim1", "SCT_snn_res.0.05"), layer = "counts")
counts_df$SCT_snn_res.0.05 <- as.factor(counts_df$SCT_snn_res.0.05)
fmla <- as.formula(paste0("Crim1", " ~ SCT_snn_res.0.05")) 
kw <- kruskal.test(fmla, data = counts_df) # statistically significant (e-12)
aggregate(Crim1 ~ SCT_snn_res.0.05, data = counts_df, median) # higher in cluster 0

counts_df <- FetchData(BNST.oxt_neurons, vars = c("Slit3", "SCT_snn_res.0.05"), layer = "counts")
counts_df$SCT_snn_res.0.05 <- as.factor(counts_df$SCT_snn_res.0.05)
fmla <- as.formula(paste0("Slit3", " ~ SCT_snn_res.0.05")) 
kw <- kruskal.test(fmla, data = counts_df) # statistically significant (e-19)
aggregate(Slit3 ~ SCT_snn_res.0.05, data = counts_df, median) # higher in cluster 1

counts_df <- FetchData(BNST.oxt_neurons, vars = c("Esr2", "SCT_snn_res.0.05"), layer = "counts")
counts_df$SCT_snn_res.0.05 <- as.factor(counts_df$SCT_snn_res.0.05)
fmla <- as.formula(paste0("Esr2", " ~ SCT_snn_res.0.05")) 
kw <- kruskal.test(fmla, data = counts_df) # not significant (0.5)
```

## Merge BNST and PVN, rerun SCT, then integrate 

```{r continue}
all_oxt_neurons_list <- list(BNST.oxt_neurons, PVN.all_oxt_neurons)
features <- SelectIntegrationFeatures(object.list = all_oxt_neurons_list, nfeatures = 3000)
all_oxt_neurons_list <- PrepSCTIntegration(object.list = all_oxt_neurons_list, anchor.features = features, assay = rep("SCT", length(all_oxt_neurons_list)))
integration_anchors <- FindIntegrationAnchors(object.list = all_oxt_neurons_list, 
                                              normalization.method = "SCT", 
                                              dims = 1:30, 
                                              anchor.features = features, 
                                              verbose = FALSE,
                                              assay = rep("SCT", length(all_oxt_neurons_list)))
all_oxt_neurons <- IntegrateData(anchorset = integration_anchors, 
                                    normalization.method = "SCT", 
                                    dims = 1:30, 
                                    verbose = FALSE)

DefaultAssay(all_oxt_neurons) <- "integrated"
all_oxt_neurons <- RunPCA(all_oxt_neurons, assay = "integrated", features = features, verbose = FALSE)
ElbowPlot(all_oxt_neurons, ndims = 50)
all_oxt_neurons <- FindNeighbors(all_oxt_neurons, assay = "integrated", reduction = "pca", dims = 1:30)
all_oxt_neurons <- FindClusters(all_oxt_neurons, resolution = 0.05)
all_oxt_neurons <- RunUMAP(all_oxt_neurons, assay = "integrated", reduction = "pca", dims = 1:30)
DimPlot(all_oxt_neurons, group.by = c("sex", "orig.ident", "condition"), ncol = 3) # mostly interspersed
```

# By-cluster DE analysis of all Oxt neurons
*See output in BNST_PVN_oxtneurons_orig_clusters_v4.pdf,* 
*all_oxt_neurons_v4_top20_markers_by_cluster.csv,*
*BNST_PVN_oxtneurons_integrated_umaps_v4.pdf*

```{r continue}
DefaultAssay(all_oxt_neurons) = "SCT"
all_oxt_neurons <- PrepSCTFindMarkers(all_oxt_neurons) 
markers_by_cluster <- FindAllMarkers(
  all_oxt_neurons,
  assay = "SCT",               
  layer = "data",              # use normalized counts in SCT assay
  only.pos = TRUE,             # return only upregulated genes 
  min.pct = 0.1,               # gene must be expressed in at least 10% of cells in either group
  logfc.threshold = 0.25       # minimum log2 fold-change
)

top20_markers <- markers_by_cluster %>%
  filter(p_val_adj < 0.05, !grepl("LOC", gene)) %>%
  group_by(cluster) %>%
  slice_max(order_by = avg_log2FC, n = 20) 
top20_markers <- top20_markers %>% arrange(as.numeric(cluster))

# reorder columns for clarity
top20_markers <- top20_markers[, c("cluster", "gene", setdiff(names(top20_markers), c("cluster", "gene")))] 

write.csv(top20_markers, file = "C:/shares/btrainor/Sinead/Fall2025/all_oxt_neurons_v4_top20_markers_by_cluster.csv", row.names = FALSE) 

all_oxt_neurons@meta.data$cluster_label <- NA
all_oxt_neurons@meta.data$cluster_label[all_oxt_neurons$SCT_snn_res.0.05 == "0"] <- "Ptgds"  # 1212 cells
all_oxt_neurons@meta.data$cluster_label[all_oxt_neurons$SCT_snn_res.0.05 == "1"] <- "Col22a1" # 307 cells
all_oxt_neurons@meta.data$cluster_label[all_oxt_neurons$SCT_snn_res.0.05 == "2"] <- "Esr2" # 108 cells
table(all_oxt_neurons$cluster_label)

DimPlot(all_oxt_neurons, reduction = "umap", group.by = "region") + ggtitle("BNST and PVN Oxt Neurons") 
# doesn't look much more interspersed, but this may just be due to small number of cells and low res
DimPlot(all_oxt_neurons_v4, reduction = "umap", group.by = "SCT_snn_res.0.05") + ggtitle("Clustering of Integrated Object") # based on clustering ran after combining PVN + BNST

table(all_oxt_neurons$SCT_snn_res.0.05, all_oxt_neurons_$region)
# cluster 0: 35% BNST
# cluster 1: 14% BNST
# cluster 2: 0% BNST

table(all_oxt_neurons$region_clusters)
saveRDS(all_oxt_neurons, file = "C:/shares/btrainor/Sinead/GitHub/all_oxt_neurons.rds") 

library(patchwork)
cluster_levels <- unique(all_oxt_neurons$region_clusters)
PVN.oxt_neurons$region_clusters  <- factor(PVN.oxt_neurons$region_clusters,  levels = cluster_levels)
BNST.oxt_neurons$region_clusters <- factor(BNST.oxt_neurons$region_clusters, levels = cluster_levels)
all_oxt_neurons$region_clusters   <- factor(all_oxt_neurons$region_clusters,   levels = cluster_levels)
my_colors <- setNames(c("#9C97CC", "#66a61e", "#d95f02", "#41b58f", "#d95b94"), cluster_levels)
cluster_sizes <- table(all_oxt_neurons$region_clusters)
all_oxt_neurons <- all_oxt_neurons[, order(cluster_sizes[all_oxt_neurons$region_clusters])] 
p1 <- DimPlot(PVN.oxt_neurons, reduction = "umap", group.by = "region_clusters", label = FALSE) +
      theme(legend.position = "none") + scale_color_manual(values = my_colors) + ggtitle("PVN")
p1 <- LabelClusters(p1, id = "region_clusters", size = 5, fontface = "bold")
b1 <- DimPlot(BNST.oxt_neurons, reduction = "umap", group.by = "region_clusters", label = FALSE) +
      theme(legend.position = "none") + scale_color_manual(values = my_colors) + ggtitle("BNST")
b1 <- LabelClusters(b1, id = "region_clusters", size = 5, fontface = "bold")
a1 <- DimPlot(all_oxt_neurons, reduction = "umap", group.by = "region_clusters", label = FALSE) +
      theme(legend.position = "none") + scale_color_manual(values = my_colors) + ggtitle("Combined") 
# this retains clustering (res0.05) from the original datasets before combining
p1 | b1 | a1
```

BNST_0 overlaps with PVN_0
BNST_1 overlaps with PVN_1

## Magno-parvo marker expression comparison across BNST and PVN
*See output in BNST_PVN_oxtneurons_integrated_featureplots_v4_sctassay.pdf*

```{r magno-parvo comparison across BNST and PVN}

magno_genes <- c("Foxp1", "Foxp2")
parvo_genes <- c("Adarb2", "Reln", "Sox5", "Esr1")

DefaultAssay(PVN.oxt_neurons) <- "SCT"
DefaultAssay(BNST.oxt_neurons) <- "SCT"
DefaultAssay(all_oxt_neurons) <- "SCT"
for (gene in c(magno_genes, parvo_genes)) {
  exprs_list <- lapply(list(BNST.oxt_neurons, PVN.oxt_neurons, all_oxt_neurons),
                       function(x) GetAssayData(x, assay = "SCT", layer = "data")[gene, ])
  vmin <- 0
  vmax <- as.numeric(quantile(unlist(exprs_list), 0.99, na.rm = TRUE))
  p1 <- FeaturePlot(PVN.oxt_neurons, features = gene, min.cutoff = vmin, max.cutoff = vmax) +
    ggtitle(paste0("PVN-", gene)) +
    theme(legend.position = "none")
  b1 <- FeaturePlot(BNST.oxt_neurons, features = gene, min.cutoff = vmin, max.cutoff = vmax) +
    ggtitle(paste0("BNST-", gene)) +
    theme(legend.position = "none")
  a1 <- FeaturePlot(all_oxt_neurons, features = gene, min.cutoff = vmin, max.cutoff = vmax) +
    ggtitle(paste0("Combined-", gene)) +
    labs(color = "Expression (SCT)") +
    theme(legend.position = "right",
          legend.title = element_text(size = 12, face = "bold", angle = 90, vjust = 0),
          legend.text = element_text(size = 10),
          legend.title.position = "right")
  combined_plot <- (p1 | b1 | a1) + plot_layout(guides = "collect")
  print(combined_plot)
}
```

## PVN marker stats
Use Kruskal-Wallis to check for significance of PVN marker expression across clusters in the integrated object.

```{r continue}
counts_df <- FetchData(all_oxt_neurons_v4, vars = c("Crim1", "SCT_snn_res.0.05"), layer = "counts")
counts_df$SCT_snn_res.0.05 <- as.factor(counts_df$SCT_snn_res.0.05)
fmla <- as.formula(paste0("Crim1", " ~ SCT_snn_res.0.05")) 
kw <- kruskal.test(fmla, data = counts_df) # statistically significant (e-47)
pairwise <- pairwise.wilcox.test(
  counts_df[["Crim1"]],
  counts_df$SCT_snn_res.0.05,
  p.adjust.method = "BH")
pairwise$p.value # cluster 0 is significantly different from both cluster 1 and 2. Clusters 1 and 2 are not sig. diff. from each other
aggregate(Crim1 ~ SCT_snn_res.0.05, data = counts_df, median) #highest in cluster 0

counts_df <- FetchData(all_oxt_neurons_v4, vars = c("Slit3", "SCT_snn_res.0.05"), layer = "counts")
counts_df$SCT_snn_res.0.05 <- as.factor(counts_df$SCT_snn_res.0.05)
fmla <- as.formula(paste0("Slit3", " ~ SCT_snn_res.0.05")) 
kw <- kruskal.test(fmla, data = counts_df) # statistically significant (e-122)
pairwise <- pairwise.wilcox.test(
  counts_df[["Slit3"]],
  counts_df$SCT_snn_res.0.05,
  p.adjust.method = "BH")
pairwise$p.value # significantly higher in cluster 1 than clusters 0 or 2. Cluster 0 is also significantly higher than cluster 2, but less so (p = 0.02)
aggregate(Slit3 ~ SCT_snn_res.0.05, data = counts_df, median) #higher in cluster 1
aggregate(Slit3 ~ SCT_snn_res.0.05, data = counts_df, mean) 

counts_df <- FetchData(all_oxt_neurons_v4, vars = c("Esr2", "SCT_snn_res.0.05"), layer = "counts")
counts_df$SCT_snn_res.0.05 <- as.factor(counts_df$SCT_snn_res.0.05)
fmla <- as.formula(paste0("Esr2", " ~ SCT_snn_res.0.05")) 
kw <- kruskal.test(fmla, data = counts_df) # significant (e-89)
pairwise <- pairwise.wilcox.test(
  counts_df[["Esr2"]],
  counts_df$SCT_snn_res.0.05,
  p.adjust.method = "BH")
pairwise$p.value # cluster 2 is significantly higher than cluster 1 or 0 as expected (because top DE). No sig. diff between clusters 0 and 1
```

