---
title: 'BNST 2025 Analysis: Oxt Neurons'
author: "Sinead Archdeacon"
date: "2026-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This section analyzes oxytocin-expressing neurons, identified during final filtering stages in the all_neurons Rmd. The BNST Oxt neurons are subclustered and further analyzed in the BNST_PVN_integration Rmd.

```{r Environment setup}

# Increase capacity of R: necessary at the start of every session
options(future.globals.maxSize = 60000 * 1024^2)

# Load libraries
library(Seurat)
library(dplyr)
library(ggplot2)
library(sctransform) 
library(harmony)
```

## Read in the data

```{r Read in data}

BNST.merged_neurons <- readRDS(file = "C:/shares/btrainor/Sinead/GitHub/BNST.merged_neurons.rds")
```

## Subsetting for Oxt neurons

Histogram to visualize oxytocin expression across all neurons:

*See output in BNST.merged_neurons_oxt_expr.pdf*

```{r find Oxt neurons}

oxt_expr <- FetchData(BNST.merged_neurons, vars = "Oxt", assay = "RNA", layer = "counts")
oxt_df <- data.frame(expression = oxt_expr$Oxt)
summary(oxt_df) #mean 0.019, 3rd quartile 0, max 71

print(ggplot(oxt_expr, aes_string(x = "Oxt")) +
        geom_histogram(binwidth = 0.5, fill = "blue", color = "black") + 
        theme_minimal() +
        labs(title = paste("Oxt Expression (RNA Counts)"),
             x = "Expression Level",
             y = "Number of Cells"))

print(ggplot(oxt_expr, aes_string(x = "Oxt")) +
        geom_histogram(binwidth = 0.5, fill = "blue", color = "black") + 
        theme_minimal() +
        xlim(1, NA) +
        labs(title = paste("Oxt Expression (RNA Counts)"),
             x = "Expression Level",
             y = "Number of Cells"))
```

I subset for Oxt+ neurons, re-normalized, and re-clustered the subset.

```{r continue}
BNST.oxt_neurons <- subset(BNST.merged_neurons, oxt_status %in% "yes") # 463 oxt neurons
table(BNST.oxt_neurons$subtype) #majority GABA (36 glut)
table(BNST.oxt_neurons$subtype, BNST.merged_oxt_neurons$legend_label) 
# cluster 9 is mostly "ambiguous" GABA/glut subtype

table(BNST.oxt_neurons$SCT_snn_res.0.3)
table(BNST.oxt_neurons$oxt_status, BNST.merged_neurons$SCT_snn_res.0.3)

# re-normalize, rerun UMAP
DefaultAssay(BNST.oxt_neurons) <- "RNA"
BNST.oxt_neurons <- NormalizeData(BNST.oxt_neurons, normalization.method = "LogNormalize", scale.factor = 10000)
BNST.oxt_neurons <- SCTransform(BNST.oxt_neurons, method = "glmGamPoi", 
                                   return.only.var.genes = FALSE, 
                                   variable.features.n = 3000, 
                                   verbose = FALSE) 
BNST.oxt_neurons <- RunPCA(BNST.oxt_neurons)
ElbowPlot(BNST.oxt_neurons, ndims =50)
BNST.oxt_neurons <- FindNeighbors(BNST.oxt_neurons, reduction = "pca", dims = 1:5) 
BNST.oxt_neurons <- FindClusters(BNST.oxt_neurons, res = 0.05)
BNST.oxt_neurons <- RunUMAP(BNST.oxt_neurons, dims = 1:5) 
table(BNST.oxt_neurons$SCT_snn_res.0.05)
table(BNST.oxt_neurons$legend_label, BNST.oxt_neurons$SCT_snn_res.0.05)
```

40 of the 43 BNST_Sim1-Oxt neurons are from Cluster 9, and all of these are from the Oxt-DE subcluster (see BNST_PVN_integration where I verified this)
However, Cluster 9 has 78 Oxt neurons so only about half of them fall into the BNST_Sim1-Oxt cluster

## Highlight Oxt neurons on all neurons to see distribution
*See output in BNST.oxt_neurons_on_all_neurons.pdf*

```{r continue}

df <- FetchData(
  BNST.cluster9_oxt,
  vars = c("clus9_subclusters", "SCT_snn_res.0.05"))
ggplot(df, aes(x = clus9_subclusters, fill = SCT_snn_res.0.05)) +
  geom_bar(position = "fill") +          # "fill" makes it proportional (0–1)
  ylab("Proportion of Cells") +
  xlab("clus9_subcluster") +
  ggtitle("Overlap of clus9_subclusters and SCT_snn_res.0.05") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal(base_size = 14)

pale_colors <- scales::hue_pal(l = 80, c = 30)(length(unique(Idents(BNST.merged_neurons_filt)))) 
p <- DimPlot(BNST.merged_neurons_filt, group.by = "legend_label", cols = pale_colors, pt.size = 0.5) + theme(legend.position = "none")
df <- FetchData(BNST.merged_neurons_filt, vars = c("umap_1", "umap_2", "legend_label")) # extract UMAP coordinates for all neurons set
df$highlight <- Cells(BNST.merged_neurons_filt) %in% Cells(BNST.merged_oxt_neurons) # highlight Oxt neurons
p + 
  geom_point(
    data = subset(df, highlight),
    aes(x = umap_1, y = umap_2),
    color = "black",
    size = 0.8,
    alpha = 0.9
  ) + ggtitle("BNST Oxt Neurons")
```

Then I inspected by-cluster Oxt expression in Oxt neurons.
*See output in BNST.oxt_neurons_oxt_expr_by_cluster.pdf, BNST.oxt_neurons_oxt_expr_by_cluster.csv*

```{r Oxt expression in Oxt neurons} 

DefaultAssay(BNST.merged_oxt_neurons) <- "RNA"
BNST.merged_oxt_neurons$SCT_snn_res.0.3 <- as.numeric(as.character(BNST.merged_oxt_neurons$SCT_snn_res.0.3))
VlnPlot(BNST.merged_oxt_neurons, features = "Oxt", group.by = "SCT_snn_res.0.3", pt.size = 0) +   
  ylab("Oxt Expression") + xlab("Cluster") + ggtitle("All Oxt-Expressing Neurons")

oxtneurons_oxtcounts <- FetchData(BNST.merged_oxt_neurons, vars = c("Oxt", "SCT_snn_res.0.3"), layer = "counts")
summary_by_cluster <- oxtneurons_oxtcounts %>%
  group_by(SCT_snn_res.0.3) %>%
  summarise(
    n_cells = n(),
    min_Oxt_reads = min(Oxt),
    mean_Oxt_reads = mean(Oxt),
    max_Oxt_reads = max(Oxt)) %>%
  arrange(as.numeric(as.character(SCT_snn_res.0.3)))

write.csv(summary_by_cluster, file = "C:/shares/btrainor/Sinead/GitHub/BNST.oxt_neurons_oxt_expr_by_cluster.csv", row.names = FALSE)
```
