---
title: "BNST 2025 Analysis: Cluster 9"
author: "Sinead Archdeacon"
output:
  html_document:
    toc: true
    toc_float: true
date: "2026-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This section focuses on analysis of cluster 9, which had the highest expression of Oxt and PVN markers of all the neuron subtype clusters.

```{r Environment setup}

# Increase capacity of R: necessary at the start of every session
options(future.globals.maxSize = 60000 * 1024^2)

# Load libraries
library(Seurat)
library(dplyr)
library(ggplot2)
library(sctransform) 
library(harmony)
```

## Read in the data

```{r Read in data}

BNST.merged_neurons <- readRDS(file = "C:/shares/btrainor/Sinead/GitHub/BNST.merged_neurons.rds")
```

## Inspect Sim1 expression
*See output in BNST.clus9_vlnplot.pdf*

```{r Sim1 expression}

# check Sim1 expression across all neurons
BNST.merged_neurons$SCT_snn_res.0.3 <- as.numeric(as.character(BNST.merged_neurons$SCT_snn_res.0.3))
DefaultAssay(BNST.merged_neurons)
VlnPlot(BNST.merged_neurons, features = "Sim1", group.by = "SCT_snn_res.0.3", pt.size = 0) + 
  ylab("Expression") + xlab("Cluster") + ggtitle("Sim1") + theme(plot.title = element_text(size = 30, face = "bold"))
```

## Subset cluster 9

```{r subset cluster 9}

BNST.neurons_cluster9 <- subset(BNST.merged_neurons, legend_label %in% "9 - Sim1 Hypothalamic-like Glut")

# re-normalize and rerun SCT on the subset
DefaultAssay(BNST.neurons_cluster9) <- "RNA"
BNST.neurons_cluster9 <- NormalizeData(BNST.neurons_cluster9, assay = "RNA") 

BNST.neurons_cluster9 <- SCTransform(BNST.neurons_cluster9, method = "glmGamPoi", 
                                    return.only.var.genes = FALSE, 
                                    variable.features.n = 3000, 
                                    verbose = FALSE) 

BNST.neurons_cluster9 <- RunPCA(BNST.neurons_cluster9, npcs = 50, verbose = FALSE) 
ElbowPlot(BNST.neurons_cluster9, ndims= 50) 
BNST.neurons_cluster9 <- FindNeighbors(BNST.neurons_cluster9, reduction = "pca", dims = 1:10) 
BNST.neurons_cluster9 <- FindClusters(BNST.neurons_cluster9, resolution = 0.12) 
BNST.neurons_cluster9 <- RunUMAP(BNST.neurons_cluster9, reduction = "pca", dims = 1:10)

DimPlot(BNST.neurons_cluster9, reduction = "umap", group.by = "SCT_snn_res.0.12")
DimPlot(BNST.neurons_cluster9, reduction = "umap", group.by = c("sex", "condition", "orig.ident"))
```

## By-cluster DE analysis for cluster 9

I ran DE analysis on just the cluster 9 subclusters, and labeled each subcluster with a biologically relevant marker gene picked from the list of top 20 marker genes.

*See output in BNST.neurons_cluster9_top20_markers_by_cluster.csv*

```{r cluster 9 subcluster by-cluster DE analysis}

DefaultAssay(BNST.neurons_cluster9) = "SCT"
BNST.neurons_cluster9 <- PrepSCTFindMarkers(BNST.neurons_cluster9) 
markers_by_cluster <- FindAllMarkers(
  BNST.neurons_cluster9,
  assay = "SCT",               
  layer = "data",              # use normalized counts in SCT assay
  only.pos = TRUE,             # return only upregulated genes 
  min.pct = 0.1,               # gene must be expressed in at least 10% of cells in either group
  logfc.threshold = 0.25       # minimum log2 fold-change
)

top20_markers <- markers_by_cluster %>%
  filter(p_val_adj < 0.05, !grepl("LOC", gene)) %>%
  group_by(cluster) %>%
  slice_max(order_by = avg_log2FC, n = 20) 
top20_markers <- top20_markers %>% arrange(as.numeric(cluster))

# reorder columns for clarity
top20_markers <- top20_markers[, c("cluster", "gene", setdiff(names(top20_markers), c("cluster", "gene")))] 

write.csv(top20_markers, file = "C:/shares/btrainor/Sinead/GitHub/BNST.neurons_cluster9_top20_markers_by_cluster.csv", row.names = FALSE) 
```

Then I labeled each cluster with a biologically relevant DE from the top 20 list.
*See output in figures/BNST.neurons_cluster9_umap_labeled.pdf*

```{r continue}
BNST.neurons_cluster9@meta.data$clus9_subclusters <- NA
BNST.neurons_cluster9@meta.data$clus9_subclusters[BNST.neurons_cluster9$SCT_snn_res.0.12 == "0"] <- "Sim1-Slit3" #442 cells
BNST.neurons_cluster9@meta.data$clus9_subclusters[BNST.neurons_cluster9$SCT_snn_res.0.12 == "1"] <- "Sim1-Bnc2" #397 cells
BNST.neurons_cluster9@meta.data$clus9_subclusters[BNST.neurons_cluster9$SCT_snn_res.0.12 == "2"] <- "Sim1-Trh" #333 cells
BNST.neurons_cluster9@meta.data$clus9_subclusters[BNST.neurons_cluster9$SCT_snn_res.0.12 == "3"] <- "Sim1-Th" #313 cells
BNST.neurons_cluster9@meta.data$clus9_subclusters[BNST.neurons_cluster9$SCT_snn_res.0.12 == "4"] <- "Sim1-Oxt" #306 cells
BNST.neurons_cluster9@meta.data$clus9_subclusters[BNST.neurons_cluster9$SCT_snn_res.0.12 == "5"] <- "Sim1-Opn5" #293 cells
BNST.neurons_cluster9@meta.data$clus9_subclusters[BNST.neurons_cluster9$SCT_snn_res.0.12 == "6"] <- "Sim1-Nxph1" #292 cells
BNST.neurons_cluster9@meta.data$clus9_subclusters[BNST.neurons_cluster9$SCT_snn_res.0.12 == "7"] <- "Sim1-Maf" #258 cells
table(BNST.neurons_cluster9$clus9_subclusters)

p <- DimPlot(BNST.neurons_cluster9, reduction = "umap", group.by = "clus9_subclusters", label = FALSE) + 
     ggtitle ("BNST Cluster 9 Neurons") + theme(legend.position = "none")
LabelClusters(p, id = "clus9_subclusters", size = 4, fontface = "bold")
```

## Visualize Oxt neuron distribution across cluster 9 subclusters.
*See output in figures/BNST.oxtneurons_on_clus9neurons.pdf*

``` {r figures}

# extract UMAP coordinates for all neurons set
df <- FetchData(BNST.neurons_cluster9, vars = c("umap_1", "umap_2", "clus9_subclusters")) 

# highlight Oxt neurons
df$highlight <- Cells(BNST.neurons_cluster9) %in% Cells(BNST.oxt_neurons) 
p + 
  geom_point(
    data = subset(df, highlight),
    aes(x = umap_1, y = umap_2),
    color = "black",
    size = 0.8,
    alpha = 0.9
  ) + ggtitle("Cluster 9 OXT Neurons")
dev.off()
highlighted_cells <- intersect(Cells(BNST.neurons_cluster9), Cells(BNST.merged_oxt_neurons))
BNST.cluster9_oxt <- subset(BNST.neurons_cluster9, cells = highlighted_cells) 
```

78 Oxt-expressing cells in cluster 9
46 cells (60%) are in the Oxt-DE cluster
40 of the Oxt-DE cells are in the BNST_Oxt-Sim1 cluster

```{r save progress}

saveRDS(BNST.neurons_cluster9, file = "C:/shares/btrainor/Sinead/GitHub/BNST.neurons_cluster9.rds")
```

