---
title: "BNST 2025 Analysis: Preprocessing"
author: "Sinead Archdeacon"
date: "2026-03-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,  
                      warning = FALSE
                      )
```

The first step of this snRNA-seq analysis is to load and pre-process the raw data from 10X Genomics CellRanger, following a typical Seurat workflow.

## Environment setup

```{r Environment setup}
# Increase capacity of R: necessary at the start of every session
options(future.globals.maxSize = 60000 * 1024^2)

# Load libraries  
library(Seurat)
library(dplyr)
library(glmGamPoi)
library(ggplot2)
library(SingleCellExperiment)
library(scDblFinder)
library(gt)
library(sctransform)
library(harmony)
```

## Seurat object creation

First, add the .h5 files to the RStudio environment.

```{r Upload files}

BNSTData_malec1 <- Read10X_h5(filename = 'C:/shares/btrainor/Sequencingdata/Resequenced_BNST/BNST_MALE_C1/filtered_feature_bc_matrix.h5', use.names = TRUE, unique.features = TRUE)
BNSTData_malec2 <- Read10X_h5(filename = 'C:/shares/btrainor/Sequencingdata/Resequenced_BNST/BNST_MALE_C2/filtered_feature_bc_matrix.h5', use.names = TRUE, unique.features = TRUE)
BNSTData_malec3 <- Read10X_h5(filename = 'C:/shares/btrainor/Sequencingdata/Resequenced_BNST/BNST_MALE_C3/filtered_feature_bc_matrix.h5', use.names = TRUE, unique.features = TRUE)
BNSTData_femc1 <- Read10X_h5(filename = 'C:/shares/btrainor/Sequencingdata/Resequenced_BNST/BNST_FEM_C1/filtered_feature_bc_matrix.h5', use.names = TRUE, unique.features = TRUE)
BNSTData_femc2 <- Read10X_h5(filename = 'C:/shares/btrainor/Sequencingdata/Resequenced_BNST/BNST_FEM_C2/filtered_feature_bc_matrix.h5', use.names = TRUE, unique.features = TRUE)
BNSTData_femc3 <- Read10X_h5(filename = 'C:/shares/btrainor/Sequencingdata/Resequenced_BNST/BNST_FEM_C3/filtered_feature_bc_matrix.h5', use.names = TRUE, unique.features = TRUE)
BNSTData_males1 <- Read10X_h5(filename = 'C:/shares/btrainor/Sequencingdata/Resequenced_BNST/BNST_MALE_S1/filtered_feature_bc_matrix.h5', use.names = TRUE, unique.features = TRUE)
BNSTData_males2 <- Read10X_h5(filename = 'C:/shares/btrainor/Sequencingdata/Resequenced_BNST/BNST_MALE_S2/filtered_feature_bc_matrix.h5', use.names = TRUE, unique.features = TRUE)
BNSTData_males3 <- Read10X_h5(filename = 'C:/shares/btrainor/Sequencingdata/Resequenced_BNST/BNST_MALE_S3/filtered_feature_bc_matrix.h5', use.names = TRUE, unique.features = TRUE)
BNSTData_fems1 <- Read10X_h5(filename = 'C:/shares/btrainor/Sequencingdata/Resequenced_BNST/BNST_FEM_S1/filtered_feature_bc_matrix.h5', use.names = TRUE, unique.features = TRUE)
BNSTData_fems2 <- Read10X_h5(filename = 'C:/shares/btrainor/Sequencingdata/Resequenced_BNST/BNST_FEM_S2/filtered_feature_bc_matrix.h5', use.names = TRUE, unique.features = TRUE)
BNSTData_fems3 <- Read10X_h5(filename = 'C:/shares/btrainor/Sequencingdata/Resequenced_BNST/BNST_FEM_S3/filtered_feature_bc_matrix.h5', use.names = TRUE, unique.features = TRUE)
```

I converted these to Seurat objects and added metadata columns for sex, condition, pooled sample identity, and percent ribosomal RNA (a helpful QC metric). You can also add percent mitochondrial RNA and Malat1 for QC, but these values were always zero in our data.

```{r Create Seurat objects}

BNST_SeuratObject_malec1 <- CreateSeuratObject(counts = BNSTData_malec1)
BNST_SeuratObject_malec2 <- CreateSeuratObject(counts = BNSTData_malec2)
BNST_SeuratObject_malec3 <- CreateSeuratObject(counts = BNSTData_malec3)
BNST_SeuratObject_femc1 <- CreateSeuratObject(counts = BNSTData_femc1)
BNST_SeuratObject_femc2 <- CreateSeuratObject(counts = BNSTData_femc2)
BNST_SeuratObject_femc3 <- CreateSeuratObject(counts = BNSTData_femc3)

BNST_SeuratObject_males1 <- CreateSeuratObject(counts = BNSTData_males1)
BNST_SeuratObject_males2 <- CreateSeuratObject(counts = BNSTData_males2)
BNST_SeuratObject_males3 <- CreateSeuratObject(counts = BNSTData_males3)
BNST_SeuratObject_fems1 <- CreateSeuratObject(counts = BNSTData_fems1)
BNST_SeuratObject_fems2 <- CreateSeuratObject(counts = BNSTData_fems2)
BNST_SeuratObject_fems3 <- CreateSeuratObject(counts = BNSTData_fems3)

# sex
BNST_SeuratObject_malec1@meta.data$sex <- "male"
BNST_SeuratObject_malec2@meta.data$sex <- "male"
BNST_SeuratObject_malec3@meta.data$sex <- "male"
BNST_SeuratObject_femc1@meta.data$sex <- "female"
BNST_SeuratObject_femc2@meta.data$sex <- "female"
BNST_SeuratObject_femc3@meta.data$sex <- "female"
BNST_SeuratObject_males1@meta.data$sex <- "male"
BNST_SeuratObject_males2@meta.data$sex <- "male"
BNST_SeuratObject_males3@meta.data$sex <- "male"
BNST_SeuratObject_fems1@meta.data$sex <- "female"
BNST_SeuratObject_fems2@meta.data$sex <- "female"
BNST_SeuratObject_fems3@meta.data$sex <- "female"

# condition
BNST_SeuratObject_malec1@meta.data$condition <- "control"
BNST_SeuratObject_malec2@meta.data$condition <- "control"
BNST_SeuratObject_malec3@meta.data$condition <- "control"
BNST_SeuratObject_femc1@meta.data$condition <- "control"
BNST_SeuratObject_femc2@meta.data$condition <- "control"
BNST_SeuratObject_femc3@meta.data$condition <- "control"
BNST_SeuratObject_males1@meta.data$condition <- "stress"
BNST_SeuratObject_males2@meta.data$condition <- "stress"
BNST_SeuratObject_males3@meta.data$condition <- "stress"
BNST_SeuratObject_fems1@meta.data$condition <- "stress"
BNST_SeuratObject_fems2@meta.data$condition <- "stress"
BNST_SeuratObject_fems3@meta.data$condition <- "stress"

# pooled sample identity
BNST_SeuratObject_malec1@meta.data$orig.ident <- "malec1"
BNST_SeuratObject_malec2@meta.data$orig.ident <- "malec2"
BNST_SeuratObject_malec3@meta.data$orig.ident <- "malec3"
BNST_SeuratObject_femc1@meta.data$orig.ident <- "femc1"
BNST_SeuratObject_femc2@meta.data$orig.ident <- "femc2"
BNST_SeuratObject_femc3@meta.data$orig.ident <- "femc3"
BNST_SeuratObject_males1@meta.data$orig.ident <- "males1"
BNST_SeuratObject_males2@meta.data$orig.ident <- "males2"
BNST_SeuratObject_males3@meta.data$orig.ident <- "males3"
BNST_SeuratObject_fems1@meta.data$orig.ident <- "fems1"
BNST_SeuratObject_fems2@meta.data$orig.ident <- "fems2"
BNST_SeuratObject_fems3@meta.data$orig.ident <- "fems3"

# rename cells to match sample identity
BNST_SeuratObject_malec1 <- RenameCells(BNST_SeuratObject_malec1, add.cell.id = "malec1_")
BNST_SeuratObject_malec2 <- RenameCells(BNST_SeuratObject_malec2, add.cell.id = "malec2_")
BNST_SeuratObject_malec3 <- RenameCells(BNST_SeuratObject_malec3, add.cell.id = "malec3_")
BNST_SeuratObject_femc1 <- RenameCells(BNST_SeuratObject_femc1, add.cell.id = "femc1_")
BNST_SeuratObject_femc2 <- RenameCells(BNST_SeuratObject_femc2, add.cell.id = "femc2_")
BNST_SeuratObject_femc3 <- RenameCells(BNST_SeuratObject_femc3, add.cell.id = "femc3_")
BNST_SeuratObject_males1 <- RenameCells(BNST_SeuratObject_males1, add.cell.id = "males1_")
BNST_SeuratObject_males2 <- RenameCells(BNST_SeuratObject_males2, add.cell.id = "males2_")
BNST_SeuratObject_males3 <- RenameCells(BNST_SeuratObject_males3, add.cell.id = "males3_")
BNST_SeuratObject_fems1 <- RenameCells(BNST_SeuratObject_fems1, add.cell.id = "fems1_")
BNST_SeuratObject_fems2 <- RenameCells(BNST_SeuratObject_fems2, add.cell.id = "fems2_")
BNST_SeuratObject_fems3 <- RenameCells(BNST_SeuratObject_fems3, add.cell.id = "fems3_")

# percent ribosomal RNA
BNST_SeuratObject_malec1[["percent.ribo"]] <- PercentageFeatureSet(BNST_SeuratObject_malec1, pattern = "^Rp[sl]")
BNST_SeuratObject_malec2[["percent.ribo"]] <- PercentageFeatureSet(BNST_SeuratObject_malec2, pattern = "^Rp[sl]")
BNST_SeuratObject_malec3[["percent.ribo"]] <- PercentageFeatureSet(BNST_SeuratObject_malec3, pattern = "^Rp[sl]")
BNST_SeuratObject_femc1[["percent.ribo"]] <- PercentageFeatureSet(BNST_SeuratObject_femc1, pattern = "^Rp[sl]")
BNST_SeuratObject_femc2[["percent.ribo"]] <- PercentageFeatureSet(BNST_SeuratObject_femc2, pattern = "^Rp[sl]")
BNST_SeuratObject_femc3[["percent.ribo"]] <- PercentageFeatureSet(BNST_SeuratObject_femc3, pattern = "^Rp[sl]")
BNST_SeuratObject_males1[["percent.ribo"]] <- PercentageFeatureSet(BNST_SeuratObject_males1, pattern = "^Rp[sl]")
BNST_SeuratObject_males2[["percent.ribo"]] <- PercentageFeatureSet(BNST_SeuratObject_males2, pattern = "^Rp[sl]")
BNST_SeuratObject_males3[["percent.ribo"]] <- PercentageFeatureSet(BNST_SeuratObject_males3, pattern = "^Rp[sl]")
BNST_SeuratObject_fems1[["percent.ribo"]] <- PercentageFeatureSet(BNST_SeuratObject_fems1, pattern = "^Rp[sl]")
BNST_SeuratObject_fems2[["percent.ribo"]] <- PercentageFeatureSet(BNST_SeuratObject_fems2, pattern = "^Rp[sl]")
BNST_SeuratObject_fems3[["percent.ribo"]] <- PercentageFeatureSet(BNST_SeuratObject_fems3, pattern = "^Rp[sl]")
```

## Quality control/filtering

Here I inspected the metrics **nFeature_RNA** (the number of unique genes in each cell) and **nCount_RNA** (the number of UMI counts per cell), which are included automatically in the CellRanger output.

A very low nFeature (below ~200-500 genes) or nCount (below ~1000 UMIs) can indicate a dead/dying cell or empty droplet (background noise). Meanwhile, a very high nFeature (above ~2500-3000 genes) or nCount (above the 95th-99th percentile of the data) can indicate a multiplet. I removed these outliers to eliminate low-quality cells from my data. 

I generated violin plots to visualize the distribution of nFeature and nCount for each sample, and scatter plots to visualize the relationship between nCount and nFeature (which should be roughly linear for high-quality cells). I inspected these plots to pick approximate thresholds for filtering. (However, more complex statistical methods can be used to generate threshold values.) Thresholds were determined separately for each object.

*See preprocessing figures for QC output*

```{r QC malec1}

# violin plot
p1 <- VlnPlot(
  BNST_SeuratObject_malec1,
  features = c("nFeature_RNA", "nCount_RNA")
) &
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) &
  xlab("malec1")
print(p1)

# scatter plot
fs1 <- FeatureScatter(BNST_SeuratObject_malec1, 
                     feature1 = "nCount_RNA", 
                     feature2 = "nFeature_RNA",
                     plot.cor = FALSE) + 
   geom_smooth(method = 'lm') +
   ggtitle("malec1") +
   theme(legend.position = "none")
print(fs1)

# filter using thresholds determined from plots
BNST_SeuratObject_malec1 <- subset(BNST_SeuratObject_malec1, 
                            subset = nFeature_RNA > 1000 & nFeature_RNA < 7500 & nCount_RNA < 40000)
```


```{r QC malec2}

# violin plot
p2 <- VlnPlot(
  BNST_SeuratObject_malec2,
  features = c("nFeature_RNA", "nCount_RNA")
) &
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) &
  xlab("malec2")
print(p2)

# scatter plot
fs2 <- FeatureScatter(BNST_SeuratObject_malec2, 
                      feature1 = "nCount_RNA", 
                      feature2 = "nFeature_RNA",
                      plot.cor = FALSE) +
   geom_smooth(method = 'lm') +
   ggtitle("malec2") +
   theme(legend.position = "none")
print(fs2)

# filter using thresholds determined from plots
BNST_SeuratObject_malec2 <- subset(BNST_SeuratObject_malec2, 
                            subset = nFeature_RNA > 1000 & nFeature_RNA < 8000 & nCount_RNA < 55000)
```


```{r QC malec3}

# violin plot
p3 <- VlnPlot(
  BNST_SeuratObject_malec3,
  features = c("nFeature_RNA", "nCount_RNA")
) &
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) &
  xlab("malec3")
print(p3)

# scatter plot
fs3 <- FeatureScatter(BNST_SeuratObject_malec3, 
                      feature1 = "nCount_RNA", 
                      feature2 = "nFeature_RNA",
                      plot.cor = FALSE) +
   geom_smooth(method = 'lm') +
   ggtitle("malec3") +
   theme(legend.position = "none")
print(fs3)

# filter using thresholds determined from plots
BNST_SeuratObject_malec3 <- subset(BNST_SeuratObject_malec3, 
                            subset = nFeature_RNA > 1000 & nFeature_RNA < 10000 & nCount_RNA < 100000)
```


```{r QC femc1}

# violin plot
p4 <- VlnPlot(
  BNST_SeuratObject_femc1,
  features = c("nFeature_RNA", "nCount_RNA")
) &
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) &
  xlab("femc1")
print(p4)

# scatter plot
fs4 <- FeatureScatter(BNST_SeuratObject_femc1, 
                      feature1 = "nCount_RNA", 
                      feature2 = "nFeature_RNA",
                      plot.cor = FALSE) + 
   geom_smooth(method = 'lm') +
   ggtitle("femc1") +
   theme(legend.position = "none")
print(fs4)

# filter using thresholds determined from plots
BNST_SeuratObject_femc1 <- subset(BNST_SeuratObject_femc1, 
                           subset = nFeature_RNA > 1000 & nFeature_RNA < 10000 & nCount_RNA < 80000)
```


```{r QC femc2}

# violin plot
p5 <- VlnPlot(
  BNST_SeuratObject_femc2,
  features = c("nFeature_RNA", "nCount_RNA")
) &
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) &
  xlab("femc2")
print(p5)

# scatter plot
fs5 <- FeatureScatter(BNST_SeuratObject_femc2, 
                      feature1 = "nCount_RNA", 
                      feature2 = "nFeature_RNA",
                      plot.cor = FALSE) + 
   geom_smooth(method = 'lm') +
   ggtitle("femc2") +
   theme(legend.position = "none")
print(fs5)

# filter using thresholds determined from plots
BNST_SeuratObject_femc2 <- subset(BNST_SeuratObject_femc2, 
                           subset = nFeature_RNA > 1000 & nFeature_RNA < 8000 & nCount_RNA < 50000)
```


```{r QC femc3}

# violin plot
p6 <- VlnPlot(
  BNST_SeuratObject_femc3,
  features = c("nFeature_RNA", "nCount_RNA")
) &
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) &
  xlab("femc3")
print(p6)

# scatter plot
fs6 <- FeatureScatter(BNST_SeuratObject_femc3, 
                      feature1 = "nCount_RNA", 
                      feature2 = "nFeature_RNA",
                      plot.cor = FALSE) +
   geom_smooth(method = 'lm') +
   ggtitle("femc3") +
   theme(legend.position = "none")
print(fs6)

# filter using thresholds determined from plots
BNST_SeuratObject_femc3 <- subset(BNST_SeuratObject_femc3, 
                           subset = nFeature_RNA > 1000 & nFeature_RNA < 11000 & nCount_RNA < 110000)
```


```{r males1}

# violin plot
p7 <- VlnPlot(
  BNST_SeuratObject_males1,
  features = c("nFeature_RNA", "nCount_RNA")
) &
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) &
  xlab("males1")
print(p7)

# scatter plot
fs7 <- FeatureScatter(BNST_SeuratObject_males1, 
                      feature1 = "nCount_RNA", 
                      feature2 = "nFeature_RNA",
                      plot.cor = FALSE) + 
   geom_smooth(method = 'lm') +
   ggtitle("males1") +
   theme(legend.position = "none")
print(fs7)

# filter using thresholds determined from plots
BNST_SeuratObject_males1 <- subset(BNST_SeuratObject_males1, 
                            subset = nFeature_RNA > 1000 & nFeature_RNA < 11000 & nCount_RNA < 80000)
```


```{r males2}

# violin plot
p8 <- VlnPlot(
  BNST_SeuratObject_males2,
  features = c("nFeature_RNA", "nCount_RNA")
) &
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) &
  xlab("males2")
print(p8)

# scatter plot
fs8 <- FeatureScatter(BNST_SeuratObject_males2, 
                      feature1 = "nCount_RNA", 
                      feature2 = "nFeature_RNA",
                      plot.cor = FALSE) + 
   geom_smooth(method = 'lm') +
   ggtitle("males2") +
   theme(legend.position = "none")
print(fs8)

# filter using thresholds determined from plots
BNST_SeuratObject_males2 <- subset(BNST_SeuratObject_males2, 
                            subset = nFeature_RNA > 1000 & nFeature_RNA < 8000 & nCount_RNA < 45000)
```


```{r males3}

# violin plot
p9 <- VlnPlot(
  BNST_SeuratObject_males3,
  features = c("nFeature_RNA", "nCount_RNA")
) &
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) &
  xlab("males3")
print(p9)

# scatter plot
fs9 <- FeatureScatter(BNST_SeuratObject_males3, 
                      feature1 = "nCount_RNA", 
                      feature2 = "nFeature_RNA",
                      plot.cor = FALSE) + 
   geom_smooth(method = 'lm') +
   ggtitle("males3") +
   theme(legend.position = "none")
print(fs9)

# filter using thresholds determined from plots
BNST_SeuratObject_males3 <- subset(BNST_SeuratObject_males3, 
                            subset = nFeature_RNA > 1000 & nFeature_RNA < 10000 & nCount_RNA < 90000)
```


```{r fems1}

# violin plot
p10 <- VlnPlot(
  BNST_SeuratObject_fems1,
  features = c("nFeature_RNA", "nCount_RNA")
) &
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) &
  xlab("fems1")
print(p10)

# scatter plot
fs10 <- FeatureScatter(BNST_SeuratObject_fems1, 
                      feature1 = "nCount_RNA", 
                      feature2 = "nFeature_RNA",
                      plot.cor = FALSE) +
   geom_smooth(method = 'lm') +
   ggtitle("fems1") +
   theme(legend.position = "none")
print(fs10)

# filter using thresholds determined from plots
BNST_SeuratObject_fems1 <- subset(BNST_SeuratObject_fems1, 
                           subset = nFeature_RNA > 1000 & nFeature_RNA < 10000 & nCount_RNA < 80000)
```


```{r fems2}

# violin plot
p11 <- VlnPlot(
  BNST_SeuratObject_fems2,
  features = c("nFeature_RNA", "nCount_RNA")
) &
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) &
  xlab("fems2")
print(p11)

# scatter plot
fs11 <- FeatureScatter(BNST_SeuratObject_fems2, 
                       feature1 = "nCount_RNA", 
                       feature2 = "nFeature_RNA",
                       plot.cor = FALSE) + 
   geom_smooth(method = 'lm') +
   ggtitle("fems2") +
   theme(legend.position = "none")
print(fs11)

# filter using thresholds determined from plots
BNST_SeuratObject_fems2 <- subset(BNST_SeuratObject_fems2, 
                           subset = nFeature_RNA > 1000 & nFeature_RNA < 8000 & nCount_RNA < 55000)
```


```{r fems3}

# violin plot
p12 <- VlnPlot(
  BNST_SeuratObject_fems3,
  features = c("nFeature_RNA", "nCount_RNA")
) &
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) &
  xlab("fems3")
print(p12)

# scatter plot
fs12 <- FeatureScatter(BNST_SeuratObject_fems3, 
                       feature1 = "nCount_RNA", 
                       feature2 = "nFeature_RNA",
                       plot.cor = FALSE) + 
   geom_smooth(method = 'lm') +
   ggtitle("fems3") +
   theme(legend.position = "none")
print(fs12)

# filter using thresholds determined from plots
BNST_SeuratObject_fems3 <- subset(BNST_SeuratObject_fems3, 
                           subset = nFeature_RNA > 1000 & nFeature_RNA < 10000 & nCount_RNA < 80000)
```

## Doublet removal

The **scDblFinder** package creates artificial doublets and trains a model to remove ambiguous cells. The model can then set threshold scores for calling doublets. I used this as a faster alternative to DoubletFinder since the Seurat objects were so large.

*See BNST_doublet_removal_summary.csv for output*

```{r Doublet removal}

# Create a list of the filtered Seurat objects
seurat_objects <- list(
  malec1 = BNST_SeuratObject_malec1,
  malec2 = BNST_SeuratObject_malec2,
  malec3 = BNST_SeuratObject_malec3,
  femc1 = BNST_SeuratObject_femc1,
  femc2 = BNST_SeuratObject_femc2,
  femc3 = BNST_SeuratObject_femc3,
  males1 = BNST_SeuratObject_males1,
  males2 = BNST_SeuratObject_males2,
  males3 = BNST_SeuratObject_males3,
  fems1 = BNST_SeuratObject_fems1,
  fems2 = BNST_SeuratObject_fems2,
  fems3 = BNST_SeuratObject_fems3
)

# Convert Seurat objects to single cell experiments (SCEs) and run scDblFinder 
sce_list <- lapply(seurat_objects, function(obj) {
  sce <- as.SingleCellExperiment(obj)
  sce <- scDblFinder(sce)  # adds scDblFinder.class + scDblFinder.score
  sce
})
```

```{r Check doublet proportions, include = FALSE}
# Check proportions of singlets vs doublets in SCEs
lapply(sce_list, function(sce) {
  prop.table(table(colData(sce)$scDblFinder.class))
})
```

```{r Continue doublet removal}
# Convert SCEs back to Seurat objects
sce_list <- lapply(
  sce_list,
  function(sce) as.Seurat(sce, counts = "counts", data = NULL)
)

# Add metadata columns for percent doublets and subset for singlets
seurat_objects_singlets <- lapply(seq_along(seurat_objects), function(i) {
  seu <- seurat_objects[[i]]
  sce <- sce_list[[i]]
  
  # extract doublet status and score
  doublet_status <- sce$scDblFinder.class
  doublet_score  <- sce$scDblFinder.score
  names(doublet_status) <- colnames(sce)
  names(doublet_score)  <- colnames(sce)
  
  # add metadata column to original Seurat object
  seu$DoubletStatus <- doublet_status[colnames(seu)]
  seu$DoubletScore  <- doublet_score[colnames(seu)]
  
  seu
})

# Inspect the doublet percentages
get_val <- function(x, name) if (name %in% names(x)) x[[name]] else 0
doublet_summary <- lapply(sce_list, function(obj) {
  tab <- table(obj$scDblFinder.class)
  
  data.frame(
    Sample = unique(obj$orig.ident)[1],
    Singlets = get_val(tab, "singlet"),
    Doublets = get_val(tab, "doublet"),
    Total = sum(tab),
    Doublet_Percent = (get_val(tab, "doublet") / sum(tab)) * 100
  )
}) %>% bind_rows()
doublet_summary %>%
  gt() %>%
  fmt_number(columns = c(Singlets, Doublets, Total), decimals = 0) %>%
  fmt_number(columns = Doublet_Percent, decimals = 2) %>%
  cols_label(
    Doublet_Percent = "Doublet (%)"
  )

# Subset singlets only
seurat_objects_singlets <- lapply(seurat_objects_singlets, function(seu) {
  subset(seu, subset = DoubletStatus == "singlet")
})
names(seurat_objects_singlets) <- names(seurat_objects)
```

```{r, include = FALSE}
write.csv(doublet_summary, "C:/shares/btrainor/Sinead/GitHub/BNST_doublet_removal_summary.csv", row.names = FALSE)
```

## Normalization and PCA

*Note:* At this point, I separated control and stress samples to avoid condition-driven effects when normalizing and clustering. The two sets of samples will be re-merged later in my analysis.

**SCTransform** normalization helps remove technical effects, such as variations in sequencing depth, by incorporating more variable features into the analysis than NormalizeData. This helps capture more subtle biological differences between cells.

The **RunPCA** function uses linear dimensionality reduction to compute orthogonal components (PCs) that explain maximum variance across cells. This is necessary prior to clustering.

**Harmony** is used to correct for batch effects among the samples.

```{r Normalization and PCA}

# Separate objects into controls and stress lists
seurat_objects_controls <- seurat_objects_singlets[c("malec1", "malec2", "malec3", "femc1", "femc2", "femc3")]
seurat_objects_stress <- seurat_objects_singlets[c("males1", "males2", "males3", "fems1", "fems2", "fems3")]

# Merge all control/stress objects into two separate objects
DefaultAssay(seurat_objects_controls[[1]]) <- "RNA"
BNST.controls <- merge(
  x = seurat_objects_controls[[1]],
  y = seurat_objects_controls[2:length(seurat_objects_controls)],
  merge.data = TRUE
)
DefaultAssay(BNST.controls) <- "RNA"
BNST.controls <- JoinLayers(object = BNST.controls, assay = "RNA", layers = "counts") 

DefaultAssay(seurat_objects_stress[[1]]) <- "RNA"
BNST.stress <- merge(
  x = seurat_objects_stress[[1]],
  y = seurat_objects_stress[2:length(seurat_objects_stress)],
  merge.data = TRUE
)
DefaultAssay(BNST.stress) <- "RNA"
BNST.stress <- JoinLayers(object = BNST.stress, assay = "RNA", layers = "counts") 

#Run SCTransform on each merged object
BNST.controls <- SCTransform(BNST.controls, method = "glmGamPoi",
                             return.only.var.genes = FALSE,
                             variable.features.n = 3000,
                             residual_type = "pearson")
BNST.stress <- SCTransform(BNST.stress, method = "glmGamPoi",
                             return.only.var.genes = FALSE,
                             variable.features.n = 3000,
                             residual_type = "pearson")

# Run PCA and batch correction (Harmony) on each object:

DefaultAssay(BNST.controls) <- "SCT"
BNST.controls <- RunPCA(BNST.controls, features = VariableFeatures(BNST.controls))
BNST.controls <- RunHarmony(BNST.controls, assay.use = "SCT", group.by.vars = "orig.ident") 

DefaultAssay(BNST.stress) <- "SCT"
BNST.stress <- RunPCA(BNST.stress, features = VariableFeatures(BNST.stress))
BNST.stress <- RunHarmony(BNST.stress, assay.use = "SCT", group.by.vars = "orig.ident") 
```

## Clustering

I used **elbow plots** to select numbers of PCs that would capture most of the variation in my data without incorporating noise. I then tested a range of clustering resolutions to find one that yielded 30-40 clusters, which will help with filtering out low-quality clusters later in my analysis. I used **RunUMAP** to compute embeddings for each cell and visualized the clusters with DimPlot.

```{r Control elbow plots}
ElbowPlot(BNST.controls, ndims = 50) +
  ggtitle("BNST control PCs")
BNST.controls <- FindNeighbors(BNST.controls, reduction = "harmony", dims = 1:40) 
```

```{r Control clustering, results = 'hide'}
res.vec <- c(0.1, 0.2, 0.3, 0.4, 0.6, 0.8, 1.0, 1.2) # vector of resolutions to test
BNST.controls <- FindClusters(BNST.controls, resolution = res.vec) # inspect the number of clusters for each resolution
BNST.controls <- FindClusters(BNST.controls, resolution = 0.5)  # gives 31 clusters, so use this res
```

```{r Control UMAPs}
set.seed(123)
BNST.controls <- RunUMAP(BNST.controls, reduction = "harmony", dims = 1:40)
DimPlot(BNST.controls, reduction = "umap", group.by = "seurat_clusters", label = TRUE) +
  ggtitle("BNST control clusters") +
  theme(legend.position = "none")
``` 

``` {r Stress elbow plots}
ElbowPlot(BNST.stress, ndims = 50) +
  ggtitle("BNST stress PCs")
BNST.stress <- FindNeighbors(BNST.stress, reduction = "harmony", dims = 1:40) 
```

```{r Stress clustering, results = 'hide'}
res.vec <- c(0.1, 0.2, 0.3, 0.4, 0.6, 0.8, 1.0, 1.2) # vector of resolutions to test
BNST.stress <- FindClusters(BNST.stress, resolution = res.vec) # inspect the number of clusters for each resolution
BNST.stress <- FindClusters(BNST.stress, resolution = 0.5)  # gives 32 clusters, so use this res
```

```{r Stress UMAPs}
set.seed(123)
BNST.stress <- RunUMAP(BNST.stress, reduction = "harmony", dims = 1:40)
DimPlot(BNST.stress, reduction = "umap", group.by = "seurat_clusters", label = TRUE) +
  ggtitle("BNST stress clusters") +
  theme(legend.position = "none")
```

## Save progress

```{r Save progress}
saveRDS(BNST.controls, file = "C:/shares/btrainor/Sinead/GitHub/BNST.controls.rds")
saveRDS(BNST.stress, file = "C:/shares/btrainor/Sinead/GitHub/BNST.stress.rds")
```
