---
title: 'BNST 2025 Analysis: All Neurons'
author: "Sinead Archdeacon"
date: "2026-03-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,  
                      warning = FALSE
                      )
```

## Introduction

The third step of this snRNA-seq analysis is to analyze all neurons and identify clusters of transcriptionally similar subtypes.

**Note:** I combined control and stress data again at this step.

```{r Environment setup}

# Increase capacity of R: necessary at the start of every session
options(future.globals.maxSize = 60000 * 1024^2)

# Load libraries
library(Seurat)
library(dplyr)
library(ggplot2)
library(sctransform) 
library(harmony)
library(presto)
```

## Control neurons

### Read in the data

```{r Read in data}

BNST.controls <- readRDS(file = "C:/shares/btrainor/Sinead/GitHub/BNST.controls_mapped.rds") # 90,729 cells
```

### Check for low-quality clusters
Check that all clusters have > 100 cells total and multiple cells from each sample. Inspect any clusters that appear very small or scattered away from the center of the UMAP.

*See output in BNST.controls_nonfilt_mapped_qc_plots.pdf*

```{r Cluster}

# cluster to get 30-40 clusters
BNST.controls_nonfilt_mapped <- FindClusters(BNST.controls_nonfilt_mapped, graph.name = "SCT_snn", resolution = 0.5) 
DimPlot(BNST.controls_nonfilt_mapped, reduction = "umap", group.by = "SCT_snn_res.0.5", label = TRUE) + ggtitle("BNST Controls UMAP") + theme(legend.position = "none")
```

### Check QC metrics again
*See output in BNST.controls_nonfilt_mapped_qc_plots.pdf*

```{r QC}

FeaturePlot(BNST.controls,
            reduction = "umap",
            features = c("nCount_RNA", "nFeature_RNA", "percent.ribo"),
            pt.size = 0.4,
            min.cutoff = "q10",
            max.cutoff = "q90",
            label = TRUE)

# make a contam_comments metadata column to see how many cells from these problematic clusters show up in neurons subset
BNST.controls_nonfilt_mapped@meta.data$contam_comments <- "NA"
BNST.controls_nonfilt_mapped@meta.data$contam_comments[BNST.controls_nonfilt_mapped$SCT_snn_res.0.5 %in% 
              c("0", "2", "6", "9", "21")] <- "low nCount" #also corresponds to low n_Feature and high percent.ribo
BNST.controls_nonfilt_mapped@meta.data$contam_comments[BNST.controls_nonfilt_mapped$SCT_snn_res.0.5 %in% 
              "4"] <- "low nCount and thalamic" #also corresponds to low n_Feature and high percent.ribo

```

### Check for contamination from neighboring brain regions

```{r check for contamination}

# dopamine receptors (striatum markers)
FeaturePlot(BNST.controls, reduction = "umap", features = c("Drd1", "Drd2"), label = TRUE) 

# olfactory tubercle marker
FeaturePlot(BNST.controls, reduction = "umap", features = "Drd3", label = TRUE)

# thalamic marker
FeaturePlot(BNST.controls, reduction = "umap", features = "Tcf7l1", label = TRUE)

# cortical/hippocampal GABAergic interneuron markers
FeaturePlot(BNST.controls, reduction = "umap", features = c("Dlx1","Dlx2"), label = TRUE)

# striatal MSN markers (Welch paper says BNST may have MSNs, so this may not be contamination)
FeaturePlot(BNST.controls, reduction = "umap", features = c("Drd1", "Drd2", "Adora2a", "Ppp1r1b"), label = TRUE)     

BNST.controls_nonfilt_mapped@meta.data$contam_comments[BNST.controls_nonfilt_mapped$SCT_snn_res.0.5 %in% 
                                                      "24"] <- "olfactory"
BNST.controls_nonfilt_mapped@meta.data$contam_comments[BNST.controls_nonfilt_mapped$SCT_snn_res.0.5 %in% 
                                                      c("10", "12", "16", "18")] <- "thalamic?"
BNST.controls_nonfilt_mapped@meta.data$contam_comments[BNST.controls_nonfilt_mapped$SCT_snn_res.0.5 %in% "8"] <- "MSN?"
```

### Try subsetting control neurons to see contamination prevalence
*See output in BNST.controls_neurons_prefilt_umap.pdf*

```{r test subset control neurons}

BNST.control_neurons_test <- subset(BNST.controls,
                      subset = (pred.div.res.0.05 %in% c("3_PAL-sAMY-TH-HY-MB-HB_neuronal", "2_Subpallium_GABAergic")))
table(BNST.control_neurons_test$contam_comments) 

# re-normalize RNA assay 
DefaultAssay(BNST.control_neurons_test) <- "RNA"
BNST.control_neurons_test <- NormalizeData(BNST.control_neurons_test, normalization.method = "LogNormalize", 
                                           scale.factor = 10000)

# Rerun SCTransform on neuron subset
BNST.control_neurons_test <- SCTransform(BNST.control_neurons_test, method = "glmGamPoi", 
                                    return.only.var.genes = FALSE, 
                                    variable.features.n = 3000, 
                                    verbose = FALSE) 

BNST.control_neurons_test <- RunPCA(BNST.control_neurons_test, npcs = 50) 
BNST.control_neurons_test <- RunHarmony(BNST.control_neurons_test, assay.use = "SCT", group.by.vars = "orig.ident")
ElbowPlot(BNST.control_neurons_test, ndims= 50)
BNST.control_neurons_test <- FindNeighbors(BNST.control_neurons_test, reduction = "harmony", dims = 1:40) 
BNST.control_neurons_test <- FindClusters(BNST.control_neurons_test, resolution = 0.3) # 22 clusters
BNST.control_neurons_test <- RunUMAP(BNST.control_neurons_test, reduction = "harmony", dims = 1:40)

DimPlot(BNST.control_neurons_test, reduction = "umap", group.by = "contam_comments", label = TRUE) 
FeaturePlot(BNST.control_neurons_test, reduction = "umap", features = "Mobp", label = TRUE) 
```

Problematic clusters account for 30,480 out of 65,723 neurons
The low nCount cluster has high Mobp expression (oligodendrocyte contamination)

### Remove low-quality or contaminant clusters

```{r filter}

BNST.controls_filt <- subset(BNST.controls, SCT_snn_res.0.5 %in% 
                            c("0", "2", "4", "6", "9", "10", "12", "16", "18", "21", "24"), invert = TRUE)
table(BNST.controls_filt$contam_comments) # check that there is nothing besides "NA" or "MSN?" remaining
# 46,096 cells remaining
```

### Subset for control neurons

```{r subset for neurons}

BNST.control_neurons <- subset(BNST.controls_filt, pred.div.res.0.05 %in% "3_PAL-sAMY-TH-HY-MB-HB_neuronal") 
# no subpallium GABAergic neurons present 
# 35,243 neurons

# re-normalize RNA assay 
DefaultAssay(BNST.control_neurons) <- "RNA"
BNST.control_neurons <- NormalizeData(BNST.control_neurons, normalization.method = "LogNormalize", scale.factor = 10000)
```

## Stress neurons

### Read in the data

```{r Read in data}

BNST.stress <- readRDS(file = "C:/shares/btrainor/Sinead/GitHub/BNST.stress_mapped.rds") # 150,378 cells
```

### Check for low-quality clusters
*See output in BNST.stress_nonfilt_mapped_qc_plots.pdf*

```{r Cluster}

# cluster to get 30-40 clusters
BNST.stress <- FindClusters(BNST.stress, graph.name = "SCT_snn", resolution = 0.5) # gives 31 clusters
DimPlot(BNST.stress, reduction = "umap", group.by = "SCT_snn_res.0.5", label = TRUE) + ggtitle("BNST Stress UMAP") + theme(legend.position = "none")
```

Cluster 30 only has 45 cells, so it can be thrown out.
Cluster 22 has only 4 cells from fems2 and 1 cell from males2 --> throw out
Cluster 23 has only 2-4 cells from 3 of the samples --> throw out
Cluster 28 has no cells from 2 of the samples --> throw out

### Check QC metrics again
*See output in BNST.stress_nonfilt_mapped_qc_plots.pdf*

```{r QC}

FeaturePlot(BNST.stress,
            reduction = "umap",
            features = c("nCount_RNA", "nFeature_RNA", "percent.ribo"),
            pt.size = 0.4,
            min.cutoff = "q10",
            max.cutoff = "q90",
            label = TRUE)

BNST.stress@meta.data$contam_comments <- "NA"
BNST.stress@meta.data$contam_comments[BNST.stress$SCT_snn_res.0.5 %in% 
                              c("0", "1", "2", "3", "5", "8")] <- "low nCount" # also low n_Feature, high percent.ribo
```

### Check for contamination from neighboring brain regions

```{r check for contamination}

# dopamine receptors (striatum markers)
FeaturePlot(BNST.stress, reduction = "umap", features = c("Drd1", "Drd2"), label = TRUE) 

# olfactory tubercle marker
FeaturePlot(BNST.stress, reduction = "umap", features = "Drd3", label = TRUE)

# thalamic marker
FeaturePlot(BNST.stress, reduction = "umap", features = "Tcf7l1", label = TRUE)

# cortical/hippocampal GABAergic interneuron markers
FeaturePlot(BNST.stress, reduction = "umap", features = c("Dlx1","Dlx2"), label = TRUE)

# striatal MSN markers 
FeaturePlot(BNST.stress, reduction = "umap", features = c("Drd1", "Drd2", "Adora2a", "Ppp1r1b"), label = TRUE) 

BNST.stress_nonfilt_mapped@meta.data$contam_comments[BNST.stress_nonfilt_mapped$SCT_snn_res.0.5 %in% 
                                    c("4", "7", "13")] <- "partial possible contam"
```

### Try subsetting stress neurons to see contamination prevalence
*See output in BNST.stress_neurons_prefilt_umap.pdf*

```{r test subset stress neurons}

BNST.stress_neurons_test <- subset(BNST.stress,
                               subset = (pred.div.res.0.02 %in% c("3_PAL-sAMY-TH-HY-MB-HB_neuronal")))
table(BNST.stress_neurons_test$contam_comments) 

# re-normalize RNA assay 
DefaultAssay(BNST.stress_neurons_test) <- "RNA"
BNST.stress_neurons_test <- NormalizeData(BNST.stress_neurons_test, normalization.method = "LogNormalize", 
                                          scale.factor = 10000)

# Rerun SCTransform on neuron subset
BNST.stress_neurons_test <- SCTransform(BNST.stress_neurons_test, method = "glmGamPoi", 
                                    return.only.var.genes = FALSE, 
                                    variable.features.n = 3000, 
                                    verbose = FALSE) 

BNST.stress_neurons_test <- RunPCA(BNST.stress_neurons_test, npcs = 50) 
BNST.stress_neurons_test <- RunHarmony(BNST.stress_neurons_test, assay.use = "SCT", group.by.vars = "orig.ident")
ElbowPlot(BNST.stress_neurons_test, ndims= 50)
BNST.stress_neurons_test <- FindNeighbors(BNST.stress_neurons_test, reduction = "harmony", dims = 1:40) 
BNST.stress_neurons_test <- FindClusters(BNST.stress_neurons_test, resolution = 0.3) #23 clusters
BNST.stress_neurons_test <- RunUMAP(BNST.stress_neurons_test, reduction = "harmony", dims = 1:40)

DimPlot(BNST.stress_neurons_test, reduction = "umap", group.by = "contam_comments", label = TRUE)
```

Problematic clusters account for 65,303 neurons out of 91,758
Clusters 0, 1, 3, 8 should be removed

### Remove low-quality or contaminant clusters

```{r filter}

BNST.stress_filt <- subset(BNST.stress, SCT_snn_res.0.5 %in% c(0, 1, 2, 3, 5, 8, 22, 23, 28, 30), invert = TRUE) 
table(BNST.stress_filt$contam_comments) # check that there is nothing besides "NA" or "partial possible contam"
# 47,326 cells remaining
```

### Subset for stress neurons

```{r subset for neurons}

BNST.stress_neurons <- subset(BNST.stress, pred.div.res.0.02 %in% "3_PAL-sAMY-TH-HY-MB-HB_neuronal") 
# 40,938 neurons

# re-normalize RNA assay 
DefaultAssay(BNST.stress_neurons) <- "RNA"
BNST.stress_neurons <- NormalizeData(BNST.stress_neurons, normalization.method = "LogNormalize", scale.factor = 10000)
```

## Merge stress and control neurons

First, I merged the subsets without integration and plotted UMAPs to check for condition-driven batch effects.

*See output in BNST.batch_effects_check.png*

```{r Merge stress and control subsets}

# add metadata column for condition 
BNST.control_neurons$condition <- "control"
BNST.stress_neurons$condition  <- "stress"

BNST.merged_neurons <- merge(BNST.control_neurons, y = BNST.stress_neurons) # 24544 neurons total

# re-run SCT
BNST.merged_neurons <- SCTransform(BNST.merged_neurons, method = "glmGamPoi", verbose = FALSE)
BNST.merged_neurons <- RunPCA(BNST.merged_neurons)
BNST.merged_neurons <- RunUMAP(BNST.merged_neurons, dims = 1:40)
DimPlot(BNST.merged_neurons, reduction = "umap", group.by = "condition") 
```

These are well-interspersed, so there are likely no condition-driven batch effects. So I can skip integration to avoid over-correction.

## By-cluster DE analysis of all neurons

Identify top differentially expressed genes for each cluster

*See output in BNST.merged_neurons_top20_markers_by_cluster.csv*

```{r by-cluster DE analysis}

DimPlot(BNST.merged_neurons, reduction = "umap", group.by = "SCT_snn_res.0.3", label = FALSE) +
  ggtitle("BNST All Neurons UMAP") + theme(legend.position = "none")

BNST.merged_neurons <- PrepSCTFindMarkers(BNST.all_neurons) 
markers_by_cluster <- FindAllMarkers(
  BNST.merged_neurons,
  assay = "SCT",               
  layer = "data",              # use normalized counts in SCT assay
  only.pos = TRUE,             # return only upregulated genes 
  min.pct = 0.1,               # gene must be expressed in at least 10% of cells in either group
  logfc.threshold = 0.25       # minimum log2 fold-change
)

top20_markers <- markers_by_cluster %>%
  filter(p_val_adj < 0.05, !grepl("LOC", gene)) %>%
  group_by(cluster) %>%
  slice_max(order_by = avg_log2FC, n = 20) 
top20_markers <- top20_markers %>% arrange(as.numeric(cluster))

# reorder columns for clarity
top20_markers <- top20_markers[, c("cluster", "gene", setdiff(names(top20_markers), c("cluster", "gene")))] 

# save markers list
write.csv(top20_markers, file = "C:/shares/btrainor/Sinead/GitHub/BNST.merged_neurons_top20_markers_by_cluster.csv", row.names = FALSE) 
```

I inspected individual clusters for predominant division and class.

```{r inspect individual clusters}

cluster0 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "0") 
table(cluster0$predicted.class) # all CNU GABA

cluster1 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "1") 
table(cluster1$predicted.class) # CNU or LSX GABA

cluster2 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "2") 
table(cluster2$predicted.class) # all CNU GABA
FeaturePlot(cluster2, features = c("Cldn14", "Ninj2", "Gli2", "Rfx4")) 
# could subcluster further and remove clusters with non-neuronal markers

cluster3 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "3") 
table(cluster3$predicted.class) # all CNU GABA
FeaturePlot(cluster3, features = "Ntsr1")

cluster4 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "4") 
table(cluster4$predicted.class) # almost all CNU GABA
FeaturePlot(cluster4, features = c("Mobp", "Mbp")) # strong expression all over, throw out this cluster
FeaturePlot(cluster4, features = "nCount_RNA") # super low

cluster5 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "5") 
table(cluster5$predicted.class) # almost all CNU GABA
FeaturePlot(cluster5, features = c("Gal", "Tac3", "Esr1")) # Tac3 and Esr1 overlap

cluster6 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "6") 
table(cluster6$predicted.class) # almost all CNU GABA
FeaturePlot(cluster6, features = c("Sst", "Nkx2-1", "Chat")) # some Nkx2-1, not much Sst or Chat

cluster7 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "7") 
table(cluster7$predicted.class) # mostly CNU-HYa Glut

cluster8 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "8") 
table(cluster8$predicted.class) # almost all CNU GABA
FeaturePlot(cluster8, features = c("Sox6", "Adarb2", "Lhx6", "Elfn1")) 
# Sox6/Lhx6/Elfn1 are MGE markers, Adarb2 is strong CGE marker, both overlap
# could subcluster

cluster9 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "9") 
table(cluster9$predicted.class) # mostly Glut
FeaturePlot(cluster9, features = c("Slc17a6", "Gad1", "Gad2", "Dlx1"))

cluster10 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "10") 
table(cluster10$predicted.class) # almost all GABA
FeaturePlot(cluster10, features = c("Tshz1", "Foxp2", "Eya2", "Dmrta2", "Tcf7l2")) # first 2 (MGE) strongly predominate

cluster11 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "11") 
table(cluster11$predicted.class) # almost all GABA

cluster12 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "12") 
table(cluster12$predicted.class) # almost all GABA

cluster13 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "13") 
table(cluster13$predicted.class) # almost all GABA
FeaturePlot(cluster13, features = c("Lhx8", "Nkx2-1", "Adarb2", "Ano1")) # overlap but Adarb2 is much stronger
cluster13 <- PrepSCTFindMarkers(cluster13)
cluster13 <- RunPCA(cluster13)
cluster13 <- FindNeighbors(cluster13, dims = 1:10)
cluster13 <- FindClusters(cluster13, res = 0.05)

cluster14 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "14") 
table(cluster14$predicted.class) # almost all GABA
table(cluster14$predicted.subclass) # mixed

cluster15 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "15") 
table(cluster15$predicted.class) # majority Glut
FeaturePlot(cluster15, features = c("Slc17a6", "Slc17a7")) # a6 prominent, a7 sparse

cluster16 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "16") 
table(cluster16$predicted.class) # almost all GABA

cluster17 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "17") 
table(cluster17$predicted.class) # majority Glut
FeaturePlot(cluster17, features = c("Slc17a6", "Slc17a7")) # a6 fairly prominent, a7 very sparse

cluster18 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "18") 
table(cluster18$predicted.class) 
# 200 MOB-DG-IMN cells--> possible olfactory bulb inhibitory neuron contamination (942 cells total)
FeaturePlot(cluster18, features = "nCount_RNA")

cluster19 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "19") 
table(cluster19$predicted.class) # almost all GABA
# QC check bc separated from rest of UMAP:
table(cluster19$orig.ident) # all samples represented
FeaturePlot(cluster19, features = c("nCount_RNA", "nFeature_RNA", "percent.ribo")) # looks alright, so keep for now

cluster20 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "20") 
table(cluster20$predicted.class) # 237 vascular, 63 olfactory bulb out of 659 total --> throw out
FeaturePlot(cluster20, features = "nCount_RNA") # low 

cluster21 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "21") 
table(cluster21$predicted.class) # almost all GABA

cluster22 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "22") 
table(cluster22$predicted.class) # all GABA

cluster23 <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% "23") 
table(cluster23$predicted.class) # almost all GABA
FeaturePlot(cluster23, features = c("Sox6", "Prox1")) # more Sox6 expression
```

I assigned Oxt identity to the merged neuron set before filtering further, to check that extra filtering wouldn't remove lots of Oxt neurons.

```{r assign Oxt identity}

DefaultAssay(BNST.merged_neurons) <- "RNA"
counts_df <- FetchData(BNST.merged_neurons, vars = c("Oxt","Avp"), layer = "counts")
oxt_neurons <- rownames(counts_df)[counts_df$Oxt >= 1 & counts_df$Oxt > counts_df$Avp] 

BNST.merged_neurons$oxt_status <- "no" 
BNST.merged_neurons$oxt_status[oxt_neurons] <- "yes"
table(BNST.merged_neurons$oxt_status) #492 (0.6%  of all neurons)
table(BNST.merged_neurons$oxt_status, BNST.merged_neurons$SCT_snn_res.0.3)
```

*Note:* Oxt >=1 gives the same result as Oxt >0 because counts can only be whole reads
Taking out the Avp parameter only adds a few more Oxt neurons, so I kept it

The Oxt neurons are spread out across all clusters, slightly enriched in cluster 9 (2.5%)
Taking out clusters 4, 18, and 20 would take out 29 oxt neurons (out of 6296 cells removed)

```{r additional round of filtering}

FeaturePlot(BNST.merged_neurons, features = c("nCount_RNA", "nFeature_RNA", "percent.ribo")) 
# clusters 4, 18, 20 account for low-quality clusters

# remove non-neuronal clusters
BNST.merged_neurons_filt <- subset(BNST.merged_neurons, SCT_snn_res.0.3 %in% c("4", "18", "20"), invert = TRUE)

# don't need to renormalize/rerun SCT bc only removed 7% of cells

BNST.merged_neurons_filt <- PrepSCTFindMarkers(BNST.merged_neurons_filt)
BNST.merged_neurons_filt <- RunPCA(BNST.merged_neurons_filt, assay = "SCT")
BNST.merged_neurons_filt <- FindNeighbors(BNST.merged_neurons_filt, dims = 1:40)
BNST.merged_neurons_filt <- RunUMAP(BNST.merged_neurons_filt, dims = 1:40)
```

Then I labeled each cluster with the most canonical/biologically relevant marker from the top 20 DEs list.

```{r label individual clusters}

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "0"] <- "MSN-like GABA" 
# Ppp1r1b, Prkch, most likely GABA

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "1"] <- "Dlx1 Prox1 CGE GABA" 

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "2"] <- "Gli2 MGE GABA"

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "3"] <- "Ntsr1 MGE GABA"

#cluster 4 thrown out (oligodendrocytes)

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "5"] <- "Esr1/2 Ar GABA"

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "6"] <- "Lhx8 Nkx2-1 MGE
GABA" # Mcm3

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "7"] <- "Lhx5 VGLUT2 Glut" 
# Slc17a6

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "8"] <- "Lhx6 Sox6 MGE GABA" 
# Sox6,Adarb2

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "9"] <- "Sim1 Hypothalamic-like Glut"  # Sim1, Sim2, Adcyap1, Onecut1 & Neurod1

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "10"] <- "Foxp2 Tshz1 MGE GABA" 

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "11"] <- "Foxp2 Nkx2-1 MGE GABA" 

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "12"] <- "Tacr3 Ngf GABA"

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "13"] <- "Adarb2 CGE GABA" 
# might be mixed

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "14"] <- "Sst MGE GABA" 
# Crhbp, Maf

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "15"] <- "Tcf7l2 Thalamic-like Glut" 

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "16"] <- "Pvalb MGE GABA" 
# Pvalb, Kcnmb4

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "17"] <- "Calb2 Glut"
# cluster 18 thrown out (MOB contamination)

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "19"] <- "Chat GABA" 
# cholinergic can be MGE or CGE

#cluster 20 thrown out (vascular) 

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "21"] <- "Crh GABA" 
# Impg1, Scn5a

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "22"] <- "VMAT2 GABA" 
# Slc18a2, Ebf1

BNST.merged_neurons$combined_seurat_clusters[BNST.merged_neurons_filt$SCT_snn_res.0.3 %in% "23"] <- "Hcrtr2 Mixed GABA" # Sox6, Prox1
```

*See output in BNST.merged_neurons_labeled3.pdf*

``` {r figure}

# create legend labels
BNST.merged_neurons$SCT_snn_res.0.3 <- as.numeric(as.character(BNST.merged_neurons$SCT_snn_res.0.3))
Idents(BNST.merged_neurons) <- "SCT_snn_res.0.3"
BNST.merged_neurons$legend_label <- paste0(
  BNST.merged_neurons$SCT_snn_res.0.3, " - ",
  BNST.merged_neurons$combined_seurat_clusters)

meta_df <- BNST.merged_neurons@meta.data
ordered_levels <- meta_df %>%
  arrange(as.numeric(SCT_snn_res.0.3)) %>%   
  pull(legend_label) %>%
  unique()
BNST.merged_neurons$legend_label <- factor(
  BNST.merged_neurons$legend_label,
  levels = ordered_levels)

# center legend labels
centers <- Embeddings(BNST.merged_neurons, "umap") %>%
  as.data.frame() %>%
  cbind(cluster = Idents(BNST.merged_neurons)) %>%
  group_by(cluster) %>%
  summarise(umap_1 = mean(umap_1), umap_2 = mean(umap_2))

DimPlot(
  BNST.merged_neurons, reduction = "umap", group.by = "legend_label", label = FALSE) + ggtitle("BNST Neuron Subtypes") +
  theme(legend.position = "right",
        legend.text = element_text(size = 10),
        legend.key.size = unit(1, "lines")) +
        guides(color = guide_legend(ncol = 1,
        override.aes = list(size = 3))) +
  geom_text(data = centers,
    aes(x = umap_1, y = umap_2, label = cluster),
    color = "black",
    size = 4)

```

``` {save progress}
saveRDS(BNST.merged_neurons, file = "C:/shares/btrainor/Sinead/GitHub/BNST.merged_neurons_labeled.rds") 
```
